---
title: 'Analyses Experiment 1: Low sampling frequency mixed dynamic range'
author: "Annelinde Lettink"
date: "3/7/2023"
output: 
  html_notebook:
    toc: true
    toc_depth: 3 # upto three depths of headings (specified by #, ##, and ###)---
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
graphics.off()

#==============================================================================
# Specify file paths and load data
shaker_experiments_folder = "/Users/annelindelettink/Documents/Work MacBook Pro Annelinde/Mechanical Shaker Machine"

datadir = paste0(shaker_experiments_folder, "/analyses/")
if (!dir.exists(datadir)) {
  stop(paste0("Directory does not exist: ", datadir))
}
# Load mechanical shaker data for experiment 1
filename_experiment = paste0(datadir, "E1_lfmr.RData") 
if (!file.exists(filename_experiment)) {
  stop(paste0("File does not exist: ", filename_experiment))
}
load(filename_experiment)

# Load functions
source("frequency_domain_functions.R")
source("time_domain_functions.R")

# Select data with low sampling frequency (25 Hz) only
index_low <- which(data$specifications$experiment == "ms_lfmr" & data$specifications$sampling_frequency == "25")
setdiff(data$specifications$label, data$specifications$label[index_low]) # Ax_215 excluded, was set to 100 Hz

data_low <- list(data = data$data[index_low], 
                  specifications = data$specifications[index_low,])
rm(data)
```

# Plot selected axis shaking direction 
```{r plot signals shaking direction lf, echo=FALSE}
plot_list <- list()
  for(signal in 1:length(data_low$data)) {
    file_name <- paste0("acceleration data: ", names(data_low$data)[signal])

    jpeg(paste0(paste0(datadir, "plots"), paste0(file_name, ".jpeg")), width=600, height=500, res=120) 

    p <- ggplot2::ggplot() + ggplot2::geom_point(data = data_low$data[[signal]], ggplot2::aes(x = time, y = SD,) , alpha = 0.2, size = 0.25) + ggplot2::theme_bw() +
      ggplot2::labs(title = file_name, y = "acceleration (g)") +
      ggplot2::scale_color_manual(values = colors)
    
    p
    dev.off()
    plot_list[[signal]] <- p
  }
```

```{r Plot signals ms_lfmr, echo =FALSE}
# Save and print the plot for each signal
for(plot in 1:length(plot_list)){
  print(plot_list[[plot]])
}
rm(plot, plot_list, p)
```
# Frequency domain analyses
```{r Settings for frequency spectra, include=FALSE}
# Plot parameters
plot = TRUE
xlabel = "Frequency (Hertz)"
ylabel = "Spectral Density (g2/Hertz)"
XLIM = c(0, 5) # Limit frequency content to 5 Hz (as 250 rpm / 60 = 4.1667 Hz is the expected max)
```

```{r Derive frequency spectra low sampling frequency experiment, echo=FALSE}
# DERIVE RAW FREQUENCY SPECTRA (for visual inspection and save data for further analysis)
raw.spectra_low <- list() 

for (file in 1:length(data_low$data)) { # for all accelerometer files
  sampling_frequency <- as.numeric(data_low$specifications$sampling_frequency[file])  
  SD <- data_low$data[[file]]$SD #axis in shaking direction
  rawSpectrum <- deriveSpectrum(SD, sampling_frequency, file_name = names(data_low$data)[[file]])
  raw.spectra_low[[file]] <- rawSpectrum
}
rm(rawSpectrum, file)
names(raw.spectra_low) <- data_low$specifications$label  

mean.freqspec_low <- deriveMeanSpectrum(raw.spectra_low)
```

```{r Save raw spectra, include=FALSE}
raw_spectra <- list(spectra = raw.spectra_low, mean_spectrum = mean.freqspec_low, specifications = data_low$specifications) # Save derived raw spectra

save(raw_spectra, file = paste0(datadir, "E1_frequency_spectra.RData"))
rm(raw.spectra_low, SD, sampling_frequency)

#load(paste0(datadir, "E1_frequency_spectra.RData"))
```

## Mean frequency spectra

```{r Plot mean frequency spectrum low sampling frequency, echo=FALSE}
plot(raw_spectra$mean_spectrum$x, raw_spectra$mean_spectrum$y, xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l", main = "Mean frequency spectrum low sampling frequency and mixed dynamic range")
rm(mean.freqspec_low)
```

## Derive frequency bins
```{r Derive frequency bins low sampling frequency, echo=FALSE}
peaks_low <- derivePeaks(raw_spectra$mean_spectrum, 25, XLIM, plot = TRUE)
bins_low <- deriveBins(peaks_low, raw_spectra$mean_spectrum, XLIM = c(0, 5), plot = TRUE) 
```

## Frequency spectra plots

### Overlay all frequency spectra
```{r Overlay all spectra low sampling frequency experiment, echo=FALSE}
plot <- plot(raw_spectra$spectra[[1]]$specx, raw_spectra$spectra[[1]]$specy, main = "Overlay all frequency spectra", 
       xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l")
for(i in 1:length(bins_low)){
    plot <- plot + abline(v=bins_low[i], col = "red")
  }  

for (spec in 2:length(raw_spectra$spectra)) {
  plot + lines(raw_spectra$spectra[[spec]]$specx, raw_spectra$spectra[[spec]]$specy, type = "l")
  
}
rm(spec, i)
```

### Seperate plots for frequency spectra 

```{r Plot spectra low sampling frequency experiment, echo=FALSE}
for (spec in 1:length(raw_spectra$spectra)) {
  spectrum <- raw_spectra$spectra[[spec]]
  title <- paste("Frequency spectrum plot: ", raw_spectra$specifications$label[[spec]])
  plot(spectrum$specx, spectrum$specy, main = title, 
       xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l")
  for(i in 1:length(bins_low)){
    abline(v=bins_low[i], col = "red")
  }  
}
rm(spec, title, spectrum, i)
```

```{r Compute dominant frequency (domFreq) and mean power spectral density (meanPSD) for the axis in the shaking direction, include = FALSE}
comparison_values <- list()
for(spectrum in 1:length(raw_spectra$spectra)){ 
  comparison_values[[spectrum]] <- deriveComparisonValues(raw_spectra$spectra[[spectrum]], bins_low)
}
names(comparison_values) <- names(raw_spectra$spectra)

#cat("\nSaving data...")
save(comparison_values, file = paste0(datadir, "E1_comparison_values.RData")) 

rm(spectrum)
```

```{r WIDE FORMAT: Prepare data set for statistical analyses, include = FALSE}
# Wide format: for both experiments and outcomes domFreq and meanPSD

df_wide <- data.frame()
id <- names(comparison_values)
df_wide <- cbind(id, data_low$specifications$brand, data_low$specifications$sampling_frequency,
                 data_low$specifications$dynamic_range)
domFreq_wide <- data.frame() 
meanPSD_wide <- data.frame()
for (accelerometer in 1:length(comparison_values)) {
  domFreq_wide <- rbind(domFreq_wide, comparison_values[[accelerometer]]$domFreq)
  meanPSD_wide <- rbind(meanPSD_wide, comparison_values[[accelerometer]]$meanPSD)
}
domFreq_ms_wide <- cbind(df_wide, domFreq_wide)
colnames(domFreq_ms_wide) <- c("serialnumber", "brand", "sampling_frequency", "dynamic_range", 
                               "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
meanPSD_ms_wide <- cbind(df_wide, meanPSD_wide)
colnames(meanPSD_ms_wide) <- c("serialnumber", "brand", "sampling_frequency", "dynamic_range", 
                               "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
rm(df_wide, id, domFreq_wide, meanPSD_wide, accelerometer)
```

```{r LONG FORMAT: Prepare data set for statistical analyses, include = FALSE}
# Combining domFreq and meanPSD (gather columns for outcome variables)
library(tidyverse)

domFreq_long <- domFreq_ms_wide %>%
  gather(key = "bin", value = "domFreq", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12") %>%
  rstatix::convert_as_factor(serialnumber, bin, brand, sampling_frequency, dynamic_range)
meanPSD_long <- meanPSD_ms_wide %>%
  gather(key = "bin", value = "meanPSD", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12") %>%
  rstatix::convert_as_factor(serialnumber, bin, brand, sampling_frequency, dynamic_range)
df_long <- merge(domFreq_long, meanPSD_long) # Join both data.frame objects to derive a single dataset
rm(domFreq_long, domFreq_ms_wide, meanPSD_long, meanPSD_ms_wide)
df_long <- df_long[order(df_long$serialnumber),]

#cat("\nSaving data...")
save(df_long, file = paste0(datadir, "E1_comparison_values_long.RData"))
rm(peaks_low, index_low)

#load(datadir, "E1_comparison_values_long.RData")
```

# Statistical comparison: Mixed model analyses

## Mean power spectral density: meanPSD ~ dynamic_range + bin + dynamic_range * bin
```{r Fit model meanPSD for low sampling frequency experiment, echo=FALSE}
meanPSD_interaction_id <- lme4::lmer(meanPSD ~ bin*dynamic_range + (1 | serialnumber), REML = TRUE, data = df_long)

# Evaluate if adding a random intercept for frequency bin is necessary
meanPSD_interaction_id_bin <- lme4::lmer(meanPSD ~ bin*dynamic_range + (1 | serialnumber) + (1 | bin), REML = TRUE, data = df_long) # No convergence, so use the simpler model
anova(meanPSD_interaction_id, meanPSD_interaction_id_bin, refit = FALSE) # This is not necessary, so use simplest model
rm(meanPSD_interaction_id_bin)
```
### Model assumption: normal distribution of residuals
```{r Mean PSD Model assumptions low sampling frequency experiment, echo=FALSE}
ggpubr::ggqqplot(df_long, "meanPSD", facet.by = "bin", title = "QQ-plot mean power spectral density")
residuals_meanPSD <- resid(meanPSD_interaction_id)
#summary(residuals_meanPSD)
hist(residuals_meanPSD)
rm(residuals_meanPSD)
```

### Model summary
```{r Model summary meanPSD low sampling frequency experiment, echo=FALSE}
car::Anova(meanPSD_interaction_id, type="II", test.statistic="F")
require(lmerTest)
test.meanPSD <- as(meanPSD_interaction_id, "merModLmerTest")
print(summary(test.meanPSD, ddf="Kenward-Roger"), correlation = FALSE)
#summary(meanPSD_interaction_id)
vcov(meanPSD_interaction_id, type = "random")
rm(test.meanPSD)
#report::report(meanPSD_interaction_id)
```

### Pairwise comparisons
#### Bin
```{r Calculate contrasts bin meanPSD for low sampling frequency experiment, echo=FALSE}
maineffect_bin <- emmeans::emmeans(meanPSD_interaction_id, comparison="pairwise", 
                                         specs= ~ bin, adjust="sidak", data = df_long)
maineffect_bin
emmeans::contrast(maineffect_bin, "pairwise")
rm(maineffect_bin)
```

# Dominant frequency: domFreq ~ dynamic_range + bin + dynamic_range*bin
```{r Fit model domFreq for low sampling frequency experiment, echo=FALSE}
domFreq_interaction_id <- lme4::lmer(domFreq ~ bin*dynamic_range + (1 | serialnumber), REML = TRUE, data = df_long)

# Evaluate if adding a random intercept for frequency bin is necessary
domFreq_interaction_id__bin <- lme4::lmer(domFreq ~ bin*dynamic_range + (1 | serialnumber) + (1 | bin), REML = TRUE, data = df_long)
anova(domFreq_interaction_id, domFreq_interaction_id__bin, refit = FALSE) # This is not necessary, so use simplest model
rm(domFreq_interaction_id__bin)
```

#### Model assumption: normal distribution of residuals
```{r Dominant frequency Model assumptions low sampling frequency experiment, echo=FALSE}
ggpubr::ggqqplot(df_long, "domFreq", facet.by = "bin", title = "QQ-plot dominant frequency")
residuals_domFreq <- resid(domFreq_interaction_id)
#summary(residuals_domFreq)
hist(residuals_domFreq) # approximately normally distributed
rm(residuals_domFreq)
```

#### Model summary
```{r Model summary domFreq low sampling frequency experiment, echo=FALSE}
car::Anova(domFreq_interaction_id, type="II", test.statistic="F")
require(lmerTest)
test.domFreq <- as(domFreq_interaction_id, "merModLmerTest")
print(summary(test.domFreq, ddf="Kenward-Roger"), correlation = FALSE)
#summary(domFreq_interaction_id)
vcov(domFreq_interaction_id, type = "random")
#report::report(domFreq_interaction_id)
rm(test.domFreq)
```

### Pairwise comparisons
#### Bin
```{r Calculate contrasts bin domFreq for low sampling frequency experiment, echo=FALSE}
maineffect_bin <- emmeans::emmeans(domFreq_interaction_id, comparison="pairwise", 
                                         specs= ~ bin, adjust="sidak", data = df_long)
maineffect_bin
emmeans::contrast(maineffect_bin, "pairwise")
rm(maineffect_bin)
```

### Boxplot
```{r Create boxplots for the experiments, include=FALSE}
bxp_psd <- createBoxplot(df_long, bins_low, outcome = "meanPSD", group = "dynamic_range") 
bxp_domFreq <- createBoxplot(df_long, bins_low, outcome = "domFreq", group = "dynamic_range")

ggplot2::ggsave(file=paste0(datadir, "plots/boxplot_E1_meanPSD.png"), bxp_psd, width = 10, height = 8, dpi = 900) #saves g
ggplot2::ggsave(file=paste0(datadir, "plots/boxplot_E1_domFreq.png"), bxp_domFreq, width = 10, height = 8, dpi = 900) #saves g
```

```{r Print boxplot for mean power spectral density, echo=FALSE}
plot(bxp_psd)
```

```{r Print boxplot for dominant frequency, echo=FALSE}
plot(bxp_domFreq)
```