---
title: 'Analyses Experiment 5: Lack of standardized orientation high sampling frequency'
author: "Annelinde Lettink"
date: "02/03/2023"
output:
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
  html_notebook:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
graphics.off()

#==============================================================================
# Specify file paths and load data
shaker_experiments_folder = "/Users/annelindelettink/Documents/Work MacBook Pro Annelinde/Mechanical Shaker Machine"

datadir = paste0(shaker_experiments_folder, "/analyses/")
if (!dir.exists(datadir)) {
  stop(paste0("Directory does not exist: ", datadir))
}
# Load mechanical shaker data for all five experiments
filename_experiment = paste0(datadir, "E5_bag.RData") 
if (!file.exists(filename_experiment)) {
  stop(paste0("File does not exist: ", filename_experiment))
}
load(filename_experiment)

# Load functions
source("frequency_domain_functions.R")
source("time_domain_functions.R")

# Select devices set to high sampling frequency of 100 Hz only
index_high <- which(round(as.numeric(data$specifications$sampling_frequency)) == "100")
data_high <- list(data = data$data[index_high], 
                  specifications = data$specifications[round(as.numeric(data$specifications$sampling_frequency)) == "100",])

setdiff(data$specifications$label, data_high$specifications$label)
rm(index_high, data)
```

# Plot acceleration signals

## Plot the selected axis shaking direction 

```{r plot signals shaking direction E5, echo=FALSE}
plot_list <- list()
  for(signal in 1:length(data_high$data)) {
    file_name <- paste0("acceleration data: ", names(data_high$data)[signal])

    jpeg(paste0(paste0(datadir, "plots"), paste0(file_name, ".jpeg")), width=600, height=500, res=120) 

    p <- ggplot2::ggplot() + ggplot2::geom_point(data = data_high$data[[signal]], ggplot2::aes(x = time, y = SD,) , alpha = 0.2, size = 0.25) + ggplot2::theme_bw() +
      ggplot2::labs(title = file_name, y = "acceleration (g)") +
      ggplot2::scale_color_manual(values = colors)
    
    p
    dev.off()
    plot_list[[signal]] <- p
  }
```

```{r Plot signals ms_bag, echo =FALSE}
# Save and print the plot for each signal
for(plot in 1:length(plot_list)){
  print(plot_list[[plot]])
}
rm(plot, plot_list, p)
```

## Plot standardised signal (acceleration devided by the standard deviation)

```{r plot list standardised signal, echo=FALSE}
plot_list <- list()
  for(signal in 1:length(data_high$data)) {
    file_name <- paste0("acceleration data: ", names(data_high$data)[signal])

    jpeg(paste0(paste0(datadir, "plots"), paste0(file_name, ".jpeg")), width=600, height=500, res=120) 
    correction <- sd(data_high$data[[signal]]$SD)
    p <- ggplot2::ggplot() + ggplot2::geom_point(data = data_high$data[[signal]], ggplot2::aes(x = time, y = SD/correction,) , alpha = 0.2, size = 0.25) + ggplot2::theme_bw() +
      ggplot2::labs(title = file_name, y = "acceleration (g)") +
      ggplot2::scale_color_manual(values = colors)
    
    p
    dev.off()
    plot_list[[signal]] <- p
  }
```

```{r Plot standardised signal , echo =FALSE}
# Save and print the plot for each signal
for(plot in 1:length(plot_list)){
  print(plot_list[[plot]])
}
rm(plot, plot_list, p)
```

## Plot standardised signal: Vector Magnitude

```{r plot list VM, echo=FALSE}
plot_list <- list()
  for(signal in 1:length(data_high$data)) {
    file_name <- paste0("acceleration data: ", names(data_high$data)[signal])

    jpeg(paste0(paste0(datadir, "plots"), paste0(file_name, ".jpeg")), width=600, height=500, res=120) 

    p <- ggplot2::ggplot() + ggplot2::geom_point(data = data_high$data[[signal]], ggplot2::aes(x = time, y = VM,) , alpha = 0.2, size = 0.25) + ggplot2::theme_bw() +
      ggplot2::labs(title = file_name, y = "acceleration (g)") +
      ggplot2::scale_color_manual(values = colors)
    
    p
    dev.off()
    plot_list[[signal]] <- p
  }
```

```{r Plot VM , echo =FALSE}
# Save and print the plot for each signal
for(plot in 1:length(plot_list)){
  print(plot_list[[plot]])
}
rm(plot, plot_list, p)
```

# Frequency spectra

```{r Settings for frequency spectra, include=FALSE}
# Plot parameters
plot = TRUE
xlabel = "Frequency (Hertz)"
ylabel = "Spectral Density (g2/Hertz)"
XLIM = c(0, 5) # Limit frequency content to 5 Hz (as 250 rpm / 60 = 4.1667 Hz is the expected max)
```

```{r Derive frequency spectra bag experiment shaker direction corrected sd and vm, echo=FALSE}
# DERIVE RAW FREQUENCY SPECTRA (for visual inspection and save data for further analysis)
raw.spectra_sd <- list() 
raw.spectra_sd_corrected <- list() 
raw.spectra_vm <- list() 

for (file in 1:length(data_high$data)) { # for all accelerometer files
  sampling_frequency <- as.numeric(data_high$specifications$sampling_frequency[file])  
  rawSpectrum_SD <- deriveSpectrum(data_high$data[[file]]$SD, sampling_frequency, 
                                   file_name = names(data_high$data)[[file]])
  rawSpectrum_VM <- deriveSpectrum(data_high$data[[file]]$VM, sampling_frequency, 
                                   file_name = names(data_high$data)[[file]])
  correction <- sd(data_high$data[[file]]$SD)
  rawSpectrum_SD_corrected <- deriveSpectrum(data_high$data[[file]]$SD/correction, sampling_frequency, 
                                             file_name = names(data_high$data)[[file]])
  raw.spectra_sd[[file]] <- rawSpectrum_SD
  raw.spectra_sd_corrected[[file]] <- rawSpectrum_SD_corrected
  raw.spectra_vm [[file]]<- rawSpectrum_VM
}
rm(rawSpectrum_SD, rawSpectrum_VM, rawSpectrum_SD_corrected, file)
names(raw.spectra_sd) <- data_high$specifications$label  
names(raw.spectra_sd_corrected) <- data_high$specifications$label  
names(raw.spectra_vm) <- data_high$specifications$label  


mean.freqspec_sd <- deriveMeanSpectrum(raw.spectra_sd)
mean.freqspec_sd_corrected <- deriveMeanSpectrum(raw.spectra_sd_corrected)
mean.freqspec_vm <- deriveMeanSpectrum(raw.spectra_vm)
```

```{r Save raw spectra, include=FALSE}
raw_spectra <- list(spectra_SD = raw.spectra_sd, mean_spectrumSD = mean.freqspec_sd, spectra_SDcorrected = raw.spectra_sd, mean_spectrumSDcorrected = mean.freqspec_sd_corrected, spectra_VM = raw.spectra_vm, mean_spectrumVM = mean.freqspec_vm, specifications = data_high$specifications) # Save derived raw spectra

save(raw_spectra, file = paste0(datadir, "E5_frequency_spectra.RData"))
rm(raw.spectra_sd, raw.spectra_sd_corrected, raw.spectra_vm, SD, sampling_frequency)

#load(paste0(datadir, "E5_frequency_spectra.RData"))
```


## Axis in the shaking direction

### Frequency spectra
#### Mean frequency spectra

```{r Plot mean frequency spectrum sd high sampling frequency, echo=FALSE}
plot(raw_spectra$mean_spectrumSD$x, raw_spectra$mean_spectrumSD$y, xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l", main = "Mean frequency spectrum (SD) high sampling frequency and mixed dynamic range")
rm(mean.freqspec_sd)
```
#### Derive frequency bins
```{r Derive frequency bins high sampling frequency, echo=FALSE}
peaks_sd <- derivePeaks(raw_spectra$mean_spectrumSD, 100, XLIM, plot = TRUE)
bins_sd <- deriveBins(peaks_sd, raw_spectra$mean_spectrumSD, XLIM = c(0, 5), plot = TRUE) 
```
#### Overlay all frequency spectra
```{r Overlay all spectra sd high sampling frequency experiment, echo=FALSE}
plot <- plot(raw_spectra$spectra_SD[[1]]$specx, raw_spectra$spectra_SD[[1]]$specy, main = "Overlay all frequency spectra", 
       xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l")
for(i in 1:length(bins_sd)){
    plot <- plot + abline(v=bins_sd[i], col = "red")
  }  

for (spec in 2:length(raw_spectra$spectra_SD)) {
  plot + lines(raw_spectra$spectra_SD[[spec]]$specx, raw_spectra$spectra_SD[[spec]]$specy, type = "l")
  
}
rm(spec, i)
```

#### Seperate plots for frequency spectra 

```{r Plot spectra sd high sampling frequency experiment, echo=FALSE}
for (spec in 1:length(raw_spectra$spectra_SD)) {
  spectrum <- raw_spectra$spectra_SD[[spec]]
  title <- paste("Frequency spectrum plot: ", raw_spectra$specifications$label[[spec]])
  plot(spectrum$specx, spectrum$specy, main = title, 
       xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l")
  for(i in 1:length(bins_sd)){
    abline(v=bins_sd[i], col = "red")
  }  
}
rm(spec, title, spectrum, i)
```
```{r Compute dominant frequency (domFreq) and mean power spectral density (meanPSD) for the axis in the shaking direction, include = FALSE}
comparison_values_sd <- list()
for(spectrum in 1:length(raw_spectra$spectra_SD)){ 
  comparison_values_sd[[spectrum]] <- deriveComparisonValues(raw_spectra$spectra_SD[[spectrum]], bins_sd)
}
names(comparison_values_sd) <- names(raw_spectra$spectra_SD)

#cat("\nSaving data...")
save(comparison_values_sd, file = paste0(datadir, "E5_comparison_values_sd.RData")) 

rm(spectrum)
```

```{r WIDE FORMAT sd: Prepare data set for statistical analyses, include = FALSE}
# Wide format: for both outcomes domFreq and meanPSD

df_wide <- data.frame()
id <- names(comparison_values_sd)
df_wide <- cbind(id, data_high$specifications$brand, data_high$specifications$sampling_frequency,
                 data_high$specifications$dynamic_range)
domFreq_wide <- data.frame() 
meanPSD_wide <- data.frame()
for (accelerometer in 1:length(comparison_values_sd)) {
  domFreq_wide <- rbind(domFreq_wide, comparison_values_sd[[accelerometer]]$domFreq)
  meanPSD_wide <- rbind(meanPSD_wide, comparison_values_sd[[accelerometer]]$meanPSD)
}
domFreq_ms_wide <- cbind(df_wide, domFreq_wide)
colnames(domFreq_ms_wide) <- c("serialnumber", "brand", "sampling_frequency", "dynamic_range", 
                               "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13")
meanPSD_ms_wide <- cbind(df_wide, meanPSD_wide)
colnames(meanPSD_ms_wide) <- c("serialnumber", "brand", "sampling_frequency", "dynamic_range", 
                               "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13")
rm(df_wide, id, domFreq_wide, meanPSD_wide, accelerometer)
```

```{r LONG FORMAT sd: Prepare data set for statistical analyses, include = FALSE}
# Combining domFreq and meanPSD (gather columns for outcome variables)
library(tidyverse)

domFreq_long <- domFreq_ms_wide %>%
  gather(key = "bin", value = "domFreq", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13") %>%
  rstatix::convert_as_factor(serialnumber, bin, brand, sampling_frequency, dynamic_range)
meanPSD_long <- meanPSD_ms_wide %>%
  gather(key = "bin", value = "meanPSD", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13") %>%
  rstatix::convert_as_factor(serialnumber, bin, brand, sampling_frequency, dynamic_range)
df_long <- merge(domFreq_long, meanPSD_long) # Join both data.frame objects to derive a single dataset
rm(domFreq_long, domFreq_ms_wide, meanPSD_long, meanPSD_ms_wide)
df_long <- df_long[order(df_long$serialnumber),]

#cat("\nSaving data...")
save(df_long, file = paste0(datadir, "E5_comparison_values_long_sd.RData"))
rm(peaks_sd)
```

#### Statistical comparison: Mixed model analyses

##### Mean power spectral density: meanPSD ~ brand + bin + brand * bin
```{r Fit model meanPSD for high sd sampling frequency experiment, echo=FALSE}
#load(paste0(datadir, "E5_comparison_values_long_sd.RData"))
meanPSD_interaction_id_sd <- lme4::lmer(meanPSD ~ bin*brand + (1 | serialnumber), REML = TRUE, data = df_long)

# Evaluate if adding a random intercept for frequency bin is necessary
meanPSD_interaction_id_bin <- lme4::lmer(meanPSD ~ bin*brand + (1 | serialnumber) + (1 | bin), REML = TRUE, data = df_long) 
anova(meanPSD_interaction_id_sd, meanPSD_interaction_id_bin, refit = FALSE) # This is not necessary, so use simplest model
rm(meanPSD_interaction_id_bin)
```

##### Model assumption: normal distribution of residuals
```{r Mean PSD Model assumptions sd high sampling frequency experiment, echo=FALSE}
ggpubr::ggqqplot(df_long, "meanPSD", facet.by = "bin", title = "QQ-plot mean power spectral density")
residuals_meanPSD_sd <- resid(meanPSD_interaction_id_sd)
#summary(residuals_meanPSD)
hist(residuals_meanPSD)
rm(residuals_meanPSD)
```

##### Model summary
```{r Model summary meanPSD sd high sampling frequency experiment, echo=FALSE}
car::Anova(meanPSD_interaction_id_sd, type="II", test.statistic="F")
require(lmerTest)
test.meanPSD <- as(meanPSD_interaction_id_sd, "merModLmerTest")
print(summary(test.meanPSD, ddf="Kenward-Roger"), correlation = FALSE)
#summary(meanPSD_interaction_id_sd)
vcov(meanPSD_interaction_id_sd, type = "random")
rm(test.meanPSD)
#report::report(meanPSD_interaction_id_sd)
```

##### Pairwise comparisons
###### Bin
```{r Calculate contrasts bin meanPSD for sd high sampling frequency experiment, echo=FALSE}
maineffect_bin <- emmeans::emmeans(meanPSD_interaction_id_sd, comparison="pairwise", 
                                         specs= ~ bin, adjust="sidak", data = df_long)
maineffect_bin
emmeans::contrast(maineffect_bin, "pairwise")
rm(maineffect_bin)
```

##### Dominant frequency: domFreq ~ dynamic_range + bin + dynamic_range*bin
```{r Fit model domFreq for sd high sampling frequency experiment, echo=FALSE}
domFreq_interaction_id_sd <- lme4::lmer(domFreq ~ bin*brand + (1 | serialnumber), REML = TRUE, data = df_long)

# Evaluate if adding a random intercept for frequency bin is necessary
domFreq_interaction_id_bin <- lme4::lmer(domFreq ~ bin*brand + (1 | serialnumber) + (1 | bin), REML = TRUE, data = df_long)
anova(domFreq_interaction_id_sd, domFreq_interaction_id_bin, refit = FALSE) # This is not necessary, so use simplest model
rm(domFreq_interaction_id_bin)
```

##### Model assumption: normal distribution of residuals
```{r Dom freq Model assumptions sd high sampling frequency experiment, echo=FALSE}
ggpubr::ggqqplot(df_long, "domFreq", facet.by = "bin", title = "QQ-plot dominant frequency spectral density")
residuals_domFreq_sd <- resid(domFreq_interaction_id_sd)
#summary(residuals_meanPSD)
hist(residuals_domFreq_sd)
rm(residuals_domFreq_sd)
```
##### Model summary
```{r Model summary domFreq sd high sampling frequency experiment, echo=FALSE}
car::Anova(domFreq_interaction_id_sd, type="II", test.statistic="F")
require(lmerTest)
test.domFreq <- as(domFreq_interaction_id_sd, "merModLmerTest")
print(summary(test.domFreq, ddf="Kenward-Roger"), correlation = FALSE)
#summary(domFreq_interaction_id_sd)
vcov(domFreq_interaction_id_sd, type = "random")
rm(test.domFreq)
#report::report(domFreq_interaction_id_sd)
```

##### Pairwise comparisons
###### Bin
```{r Calculate contrasts bin domFreq for sd high sampling frequency experiment, echo=FALSE}
maineffect_bin <- emmeans::emmeans(domFreq_interaction_id_sd, comparison="pairwise", 
                                         specs= ~ bin, adjust="sidak", data = df_long)
maineffect_bin
emmeans::contrast(maineffect_bin, "pairwise")
rm(maineffect_bin)
```

###### Brand
```{r Calculate contrasts brand domFreq for sd high sampling frequency experiment, echo=FALSE}
maineffect_brand <- emmeans::emmeans(domFreq_interaction_id_sd, comparison="pairwise", 
                                         specs= ~ brand, adjust="sidak", data = df_long)
maineffect_brand
emmeans::contrast(maineffect_brand, "pairwise")
rm(maineffect_brand)
```

###### Interaction effect
```{r Calculate contrasts interaction effect domFreq for sd high sampling frequency experiment, echo=FALSE}
interactioneffect <- emmeans::emmeans(domFreq_interaction_id_sd, comparison="pairwise", 
                                         specs= ~ bin * brand, adjust="sidak", data = df_long)

ct <- emmeans::contrast(interactioneffect, "pairwise")
ct.df <- as.data.frame(ct)

focusOfInterest <- c()
  check <- strsplit(ct.df$contrast, " ")
  for (row in 1:nrow(ct.df)) {
    focusOfInterest <- c(focusOfInterest, (check[[row]][1] == check[[row]][4] 
                                           & check[[row]][2] !=   check[[row]][5])) #focus on rows where bin is the same, but the brand differs.
  }
ct.df[focusOfInterest,]
rm(ct, interactioneffect)

```

##### Boxplot
```{r Create boxplots for the sd, include=FALSE}
bxp_psd <- createBoxplot(df_long, bins_sd, outcome = "meanPSD", group = "brand", sampling_frequency = "E5") 
bxp_domFreq <- createBoxplot(df_long, bins_sd, outcome = "domFreq", group = "brand", sampling_frequency = "E5")

ggplot2::ggsave(file=paste0(datadir, "plots/boxplot_E5_meanPSD_sd.png"), bxp_psd, width = 10, height = 8, dpi = 900) #saves g
ggplot2::ggsave(file=paste0(datadir, "plots/boxplot_E5_domFreq_sd.png"), bxp_domFreq, width = 10, height = 8, dpi = 900) #saves g
```

```{r Print boxplot for sd mean power spectral density, echo=FALSE}
plot(bxp_psd)
```

```{r Print boxplot for sd dominant frequency, echo=FALSE}
plot(bxp_psd)
```

### Cross-correlation coefficients

```{r Compute cross-correlation matrices sd, include=FALSE}
correlation_matrices <- computeCrossCorrelations(data_high$data, plot = FALSE) 
save(correlation_matrices, file = paste0(datadir, "E5_cross_correlations_overall_sd.RData"))

correlations_perShakerCondition <- computeCrossCorrelationsPerShakerCondition(data_high$data) 
save(correlations_perShakerCondition, file = paste0(datadir, "E5_cross_correlations_perShakerCondition_sd.RData"))

#load(paste0(datadir, "E5_cross_correlations_overall_sd.RData"))
#load(paste0(datadir, "E5_cross_correlations_perShakerCondition_sd.RData"))
```

#### Overall heatmap
```{r Heatmap sd high sampling frequency, echo=FALSE, fig.width=14, fig.height=14}
fullmatrix <- correlation_matrices$correlationMatrix

heatmap <- corrplot::corrplot(pmax(fullmatrix, t(fullmatrix), na.rm = TRUE), 
                                  type = 'lower', method = 'color', order = "alphabet", 
                                  number.cex = 1.25, cl.cex = 2, tl.cex = 2,
                                  col.lim = c(0,1), addCoef.col = 'black', tl.srt = 45, 
                                  diag = TRUE, tl.col = "black")
rm(fullmatrix, heatmap)
```

#### Heatmap per shaker condition
```{r Heatmap sd high sampling frequency per shaker condition, echo=FALSE, fig.width=14, fig.height=15}
fullmatrices <- correlations_perShakerCondition

heatmaps <- list()
for (condition in 1:length(fullmatrices)){
  heatmaps[[condition]] <- corrplot::corrplot(pmax(fullmatrices[[condition]]$correlationMatrix,
                          t(fullmatrices[[condition]]$correlationMatrix), na.rm = TRUE), 
                            title = paste0("shaker condition: ", names(fullmatrices)[condition]),
                            type = 'lower', method = 'color', order = "alphabet", 
                            number.cex = 1.25, cl.cex = 2, tl.cex = 2,
                            col.lim = c(-1,1), addCoef.col = 'black', tl.srt = 45, mar=c(0,0,2,0),
                            diag = TRUE, tl.col = "black")
}

rm(fullmatrices, heatmaps)
```

### Comparison of cross-correlation coefficients between brands

```{r Calculate average maximum cross lagged correlation coefficients sd high sampling frequency per shaker condition, include=FALSE}
avg_crosscor_between <- list()
avg_crosscor_within <- list()

for (condition in 1:length(correlations_perShakerCondition)){
  avg_crosscor <- dataframeShapeComparisonBeween(correlations_perShakerCondition[[condition]], data_high$specifications, between = "brand") 
  avg_crosscor_between[[condition]] <- avg_crosscor[avg_crosscor$same_group == "0",]
  avg_crosscor_within[[condition]] <- avg_crosscor[avg_crosscor$same_group == "1",]
}
rm(avg_crosscor, condition)
names(avg_crosscor_between) <- names(correlations_perShakerCondition)
names(avg_crosscor_within) <- names(correlations_perShakerCondition)
```

```{r Between brand comparison sd high sampling frequency, echo=FALSE}
table_between <- list()

for(condition in 1:length(avg_crosscor_between)){
  df_correlations <- avg_crosscor_between[[condition]]
  df_correlations$correlation <- as.numeric(df_correlations$correlation)
  avg_cor <- tapply(df_correlations$correlation, df_correlations$between, summary)
  SD <- tapply(df_correlations$correlation, df_correlations$between, sd)
  sd <- c(SD[2], SD[4], SD[5], SD[7], SD[8], SD[9])
  tab <- rbind(avg_cor$AxivityActigraph, avg_cor$GENEActivActigraph, avg_cor$GENEActivAxivity, 
               avg_cor$MOXActigraph, avg_cor$MOXAxivity, avg_cor$MOXGENEActiv)
  tab <- cbind(tab, sd)
  tab <- tab[,-1]
  tab <- tab[,-5]
  table_between[[condition]] <- tab
}
rm(avg_cor, tab, SD, sd, condition)
names(table_between) <- names(correlations_perShakerCondition)
```

```{r Print between comparison tables sd high frequency, echo = FALSE}
for(condition in 1:length(table_between)){
  print(paste0("Shaker condition: ", names(table_between)[condition]))
  print(table_between[[condition]])
}
```

```{r Line graph high sampling frequency, echo=FALSE}
shaker_frequencies <- c(30, 40, 50, 62, 75, 87, 100, 112, 125, 137, 150, 160, 175, 187, 200, 212, 225, 237, 250)
groups <- rownames(table_between[[1]])
comparison <- rep(groups, length(shaker_frequencies))
correlations <- c()
se <- c()
shaker_frequency <- c()

for(condition in 1:length(table_between)){
  correlations <- c(correlations, table_between[[condition]][,3])
  se <- c(se, table_between[[condition]][,5]*1.96)
  shaker_frequency <- c(shaker_frequency, rep(shaker_frequencies[condition], length(table_between[[condition]][,3])))
}
overview_tab_between <- cbind(comparison, correlations, shaker_frequency, se)
overview_tab_between <- as.data.frame(overview_tab_between)

line_plot <- ggplot2::ggplot(data = overview_tab_between, ggplot2::aes(x = as.factor(shaker_frequency), y = as.numeric(correlations), group = as.factor(comparison))) +
  ggplot2::geom_line(ggplot2::aes(color = as.factor(comparison), linetype = as.factor(comparison))) +
  ggplot2::geom_point(ggplot2::aes(color = as.factor(comparison), shape = as.factor(comparison))) +
  ggplot2::labs(x="Shaker frequency (rpm)", y = "Cross-correlation coefficient", color = "Brand comparison") +
  ggplot2::scale_x_discrete(limits = c("30", "40", "50", "62", "75", "87", "100", "112", "125", "137", "150", "160", "175", "187", "200", "212", "225", "237", "250")) +
  scale_color_manual(values = c("orange", "green", "blue", "red", "purple", "yellow")) +
  ggplot2::theme_classic() +
  ggplot2::theme(legend.position="bottom") +
  ggplot2::geom_hline(yintercept=0.25, linetype="dashed") +
  ggplot2::geom_hline(yintercept=0.5, linetype="dashed") +
  ggplot2::geom_hline(yintercept=0.75, linetype="dashed") +
  
  ggplot2::geom_errorbar(ggplot2::aes(ymin = as.numeric(correlations) - as.numeric(se),
                                      ymax = as.numeric(correlations) + as.numeric(se), 
                                      color = as.factor(comparison)), 
                         width=0.1) + 
    ggplot2::labs(color  = "Brand comparison", linetype = "Brand comparison", shape = "Brand comparison")


ggplot2::ggsave(file=paste0(datadir, "plots/linegraph_E5_crosscorrelations_sd.png"), line_plot, width = 10, height = 8, dpi = 900) #saves g
```

```{r Plot line graph sd, echo=TRUE}
plot(line_plot)
```