---
title: 'Analyses Experiment 4: High sampling frequency mixed dynamic range'
author: "Annelinde Lettink"
date: "3/7/2023"
output:
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
  html_notebook:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
graphics.off()

#==============================================================================
# Specify file paths and load data
shaker_experiments_folder = "/Users/annelindelettink/Documents/Work MacBook Pro Annelinde/Mechanical Shaker Machine"

datadir = paste0(shaker_experiments_folder, "/analyses/")
if (!dir.exists(datadir)) {
  stop(paste0("Directory does not exist: ", datadir))
}
# Load mechanical shaker data for experiment 4
filename_experiment = paste0(datadir, "E4_hfmr.RData") 
if (!file.exists(filename_experiment)) {
  stop(paste0("File does not exist: ", filename_experiment))
}
load(filename_experiment)

# Load functions
source("frequency_domain_functions.R")
source("time_domain_functions.R")

# Select data with low sampling frequency (100 Hz) only
index_high <- which(data$specifications$experiment == "ms_hfmr" & data$specifications$sampling_frequency == "100")

data_high <- list(data = data$data[index_high], 
                  specifications = data$specifications[index_high,])
rm(data)
```

# Plot selected axis shaking direction 
```{r plot signals shaking direction hf, echo=FALSE}
plot_list <- list()
  for(signal in 1:length(data_high$data)) {
    file_name <- paste0("acceleration data: ", names(data_high$data)[signal])

    jpeg(paste0(paste0(datadir, "plots"), paste0(file_name, ".jpeg")), width=600, height=500, res=120) 

    p <- ggplot2::ggplot() + ggplot2::geom_point(data = data_high$data[[signal]], ggplot2::aes(x = time, y = SD,) , alpha = 0.2, size = 0.25) + ggplot2::theme_bw() +
      ggplot2::labs(title = file_name, y = "acceleration (g)") +
      ggplot2::scale_color_manual(values = colors)
    
    p
    dev.off()
    plot_list[[signal]] <- p
  }
```

```{r Plot signals ms_hfmr, echo =FALSE}
# Save and print the plot for each signal
for(plot in 1:length(plot_list)){
  print(plot_list[[plot]])
}
rm(plot, plot_list, p)
```
# Frequency domain analyses

```{r Settings for frequency spectra, include=FALSE}
# Plot parameters
plot = TRUE
xlabel = "Frequency (Hertz)"
ylabel = "Spectral Density (g2/Hertz)"
XLIM = c(0, 5) # Limit frequency content to 5 Hz (as 250 rpm / 60 = 4.1667 Hz is the expected max)
```

```{r Derive frequency spectra high sampling frequency experiment, echo=FALSE}
# DERIVE RAW FREQUENCY SPECTRA (for visual inspection and save data for further analysis)
raw.spectra_high <- list() 

for (file in 1:length(data_high$data)) { # for all accelerometer files
  sampling_frequency <- as.numeric(data_high$specifications$sampling_frequency[file])  
  SD <- data_high$data[[file]]$SD #axis in shaking direction
  rawSpectrum <- deriveSpectrum(SD, sampling_frequency, file_name = names(data_high$data)[[file]])
  raw.spectra_high[[file]] <- rawSpectrum
}
rm(rawSpectrum, file)
names(raw.spectra_high) <- data_high$specifications$label  

mean.freqspec_high <- deriveMeanSpectrum(raw.spectra_high)
```

```{r Save raw spectra, include=FALSE}
raw_spectra <- list(spectra = raw.spectra_high, mean_spectrum = mean.freqspec_high, specifications = data_high$specifications) # Save derived raw spectra

save(raw_spectra, file = paste0(datadir, "E4_frequency_spectra.RData"))
rm(raw.spectra_high, SD, sampling_frequency)

#load(paste0(datadir, "E4_frequency_spectra.RData"))
```

## Mean frequency spectra

```{r Plot mean frequency spectrum high sampling frequency, echo=FALSE}
plot(raw_spectra$mean_spectrum$x, raw_spectra$mean_spectrum$y, xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l", main = "Mean frequency spectrum high sampling frequency and mixed dynamic range")
rm(mean.freqspec_high)
```

## Derive frequency bins
```{r Derive frequency bins high sampling frequency, echo=FALSE}
peaks_high <- derivePeaks(raw_spectra$mean_spectrum, 100, XLIM, plot = TRUE)
bins_high <- deriveBins(peaks_high, raw_spectra$mean_spectrum, XLIM = c(0, 5), plot = TRUE) 
```

## Frequency spectra plots

### Overlay all frequency spectra
```{r Overlay all spectra high sampling frequency experiment, echo=FALSE}
plot <- plot(raw_spectra$spectra[[1]]$specx, raw_spectra$spectra[[1]]$specy, main = "Overlay all frequency spectra", 
       xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l")
for(i in 1:length(bins_high)){
    plot <- plot + abline(v=bins_high[i], col = "red")
  }  

for (spec in 2:length(raw_spectra$spectra)) {
  plot + lines(raw_spectra$spectra[[spec]]$specx, raw_spectra$spectra[[spec]]$specy, type = "l")
  
}
rm(spec, i)
```

### Seperate plots for frequency spectra 

```{r Plot spectra high sampling frequency experiment, echo=FALSE}
for (spec in 1:length(raw_spectra$spectra)) {
  spectrum <- raw_spectra$spectra[[spec]]
  title <- paste("Frequency spectrum plot: ", raw_spectra$specifications$label[[spec]])
  plot(spectrum$specx, spectrum$specy, main = title, 
       xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l")
  for(i in 1:length(bins_high)){
    abline(v=bins_high[i], col = "red")
  }  
}
rm(spec, title, spectrum, i)
```

```{r Compute dominant frequency (domFreq) and mean power spectral density (meanPSD) for the axis in the shaking direction, include = FALSE}
comparison_values <- list()
for(spectrum in 1:length(raw_spectra$spectra)){ 
  comparison_values[[spectrum]] <- deriveComparisonValues(raw_spectra$spectra[[spectrum]], bins_high)
}
names(comparison_values) <- names(raw_spectra$spectra)

#cat("\nSaving data...")
save(comparison_values, file = paste0(datadir, "E4_comparison_values.RData")) 

rm(spectrum)
```

```{r WIDE FORMAT: Prepare data set for statistical analyses, include = FALSE}
# Wide format: for both outcomes domFreq and meanPSD

df_wide <- data.frame()
id <- names(comparison_values)
df_wide <- cbind(id, data_high$specifications$brand, data_high$specifications$sampling_frequency,
                 data_high$specifications$dynamic_range)
domFreq_wide <- data.frame() 
meanPSD_wide <- data.frame()
for (accelerometer in 1:length(comparison_values)) {
  domFreq_wide <- rbind(domFreq_wide, comparison_values[[accelerometer]]$domFreq)
  meanPSD_wide <- rbind(meanPSD_wide, comparison_values[[accelerometer]]$meanPSD)
}
domFreq_ms_wide <- cbind(df_wide, domFreq_wide)
colnames(domFreq_ms_wide) <- c("serialnumber", "brand", "sampling_frequency", "dynamic_range", 
                               "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14")
meanPSD_ms_wide <- cbind(df_wide, meanPSD_wide)
colnames(meanPSD_ms_wide) <- c("serialnumber", "brand", "sampling_frequency", "dynamic_range", 
                               "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14")
rm(df_wide, id, domFreq_wide, meanPSD_wide, accelerometer)
```

```{r LONG FORMAT: Prepare data set for statistical analyses, include = FALSE}
# Combining domFreq and meanPSD (gather columns for outcome variables)
library(tidyverse)

domFreq_long <- domFreq_ms_wide %>%
  gather(key = "bin", value = "domFreq", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14") %>%
  rstatix::convert_as_factor(serialnumber, bin, brand, sampling_frequency, dynamic_range)
meanPSD_long <- meanPSD_ms_wide %>%
  gather(key = "bin", value = "meanPSD", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14") %>%
  rstatix::convert_as_factor(serialnumber, bin, brand, sampling_frequency, dynamic_range)
df_long <- merge(domFreq_long, meanPSD_long) # Join both data.frame objects to derive a single dataset
rm(domFreq_long, domFreq_ms_wide, meanPSD_long, meanPSD_ms_wide)
df_long <- df_long[order(df_long$serialnumber),]

#cat("\nSaving data...")
save(df_long, file = paste0(datadir, "E4_comparison_values_long.RData"))
rm(peaks_high, index_high)

#load(datadir, "E4_comparison_values_long.RData")
```

# Statistical comparison: Mixed model analyses

## Mean power spectral density: meanPSD ~ dynamic_range + bin + dynamic_range * bin
```{r Fit model meanPSD for low sampling frequency experiment, echo=FALSE}
meanPSD_interaction_id <- lme4::lmer(meanPSD ~ bin*dynamic_range + (1 | serialnumber), REML = TRUE, data = df_long)

# Evaluate if adding a random intercept for frequency bin is necessary
meanPSD_interaction_id_bin <- lme4::lmer(meanPSD ~ bin*dynamic_range + (1 | serialnumber) + (1 | bin), REML = TRUE, data = df_long) # No convergence, so use the simpler model
anova(meanPSD_interaction_id, meanPSD_interaction_id_bin, refit = FALSE) # This is not necessary, so use simplest model
rm(meanPSD_interaction_id_bin)
```
### Model assumption: normal distribution of residuals
```{r Mean PSD Model assumptions low sampling frequency experiment, echo=FALSE}
ggpubr::ggqqplot(df_long, "meanPSD", facet.by = "bin", title = "QQ-plot mean power spectral density")
residuals_meanPSD <- resid(meanPSD_interaction_id)
#summary(residuals_meanPSD)
hist(residuals_meanPSD)
rm(residuals_meanPSD)
```

### Model summary
```{r Model summary meanPSD low sampling frequency experiment, echo=FALSE}
car::Anova(meanPSD_interaction_id, type="II", test.statistic="F")
require(lmerTest)
test.meanPSD <- as(meanPSD_interaction_id, "merModLmerTest")
print(summary(test.meanPSD, ddf="Kenward-Roger"), correlation = FALSE)
#summary(meanPSD_interaction_id)
vcov(meanPSD_interaction_id, type = "random")
rm(test.meanPSD)
#report::report(meanPSD_interaction_id)
```

### Pairwise comparisons
#### Bin
```{r Calculate contrasts bin meanPSD for low sampling frequency experiment, echo=FALSE}
maineffect_bin <- emmeans::emmeans(meanPSD_interaction_id, comparison="pairwise", 
                                         specs= ~ bin, adjust="sidak", data = df_long)
maineffect_bin
emmeans::contrast(maineffect_bin, "pairwise")
rm(maineffect_bin)
```

# Dominant frequency: domFreq ~ dynamic_range + bin + dynamic_range*bin
```{r Fit model domFreq for low sampling frequency experiment, echo=FALSE}
domFreq_interaction_id <- lme4::lmer(domFreq ~ bin*dynamic_range + (1 | serialnumber), REML = TRUE, data = df_long)

# Evaluate if adding a random intercept for frequency bin is necessary
domFreq_interaction_id__bin <- lme4::lmer(domFreq ~ bin*dynamic_range + (1 | serialnumber) + (1 | bin), REML = TRUE, data = df_long)
anova(domFreq_interaction_id, domFreq_interaction_id__bin, refit = FALSE) # This is not necessary, so use simplest model
rm(domFreq_interaction_id__bin)
```

#### Model assumption: normal distribution of residuals
```{r Dominant frequency Model assumptions low sampling frequency experiment, echo=FALSE}
ggpubr::ggqqplot(df_long, "domFreq", facet.by = "bin", title = "QQ-plot dominant frequency")
residuals_domFreq <- resid(domFreq_interaction_id)
#summary(residuals_domFreq)
hist(residuals_domFreq) # approximately normally distributed
rm(residuals_domFreq)
```
#### Model summary
```{r Model summary domFreq low sampling frequency experiment, echo=FALSE}
car::Anova(domFreq_interaction_id, type="II", test.statistic="F")
require(lmerTest)
test.domFreq <- as(domFreq_interaction_id, "merModLmerTest")
print(summary(test.domFreq, ddf="Kenward-Roger"), correlation = FALSE)
#summary(domFreq_interaction_id)
vcov(domFreq_interaction_id, type = "random")
#report::report(domFreq_interaction_id)
rm(test.domFreq)
```

### Pairwise comparisons
#### Bin
```{r Calculate contrasts bin domFreq for low sampling frequency experiment, echo=FALSE}
maineffect_bin <- emmeans::emmeans(domFreq_interaction_id, comparison="pairwise", 
                                         specs= ~ bin, adjust="sidak", data = df_long)
maineffect_bin
emmeans::contrast(maineffect_bin, "pairwise")
rm(maineffect_bin)
```

### Boxplot
```{r Create boxplots for the experiments, include=FALSE}
bxp_psd <- createBoxplot(df_long, bins_high, outcome = "meanPSD", group = "dynamic_range") 
bxp_domFreq <- createBoxplot(df_long, bins_high, outcome = "domFreq", group = "dynamic_range")

ggplot2::ggsave(file=paste0(datadir, "plots/boxplot_E4_meanPSD.png"), bxp_psd, width = 10, height = 8, dpi = 900) #saves g
ggplot2::ggsave(file=paste0(datadir, "plots/boxplot_E4_domFreq.png"), bxp_domFreq, width = 10, height = 8, dpi = 900) #saves g
```

```{r Print boxplot for mean power spectral density, echo=FALSE}
plot(bxp_psd)
```

```{r Print boxplot for dominant frequency, echo=FALSE}
plot(bxp_domFreq)
```
# Time domain analyses

## Cross-correlation coefficients

```{r Compute cross-correlation matrices, include=FALSE}
correlation_matrices <- computeCrossCorrelations(data_high$data, plot = FALSE) 
save(correlation_matrices, file = paste0(datadir, "E4_cross_correlations_overall.RData"))

correlations_perShakerCondition <- computeCrossCorrelationsPerShakerCondition(data_high$data) 
save(correlations_perShakerCondition, file = paste0(datadir, "E4_cross_correlations_perShakerCondition.RData"))
```

### Overall heatmap
```{r Heatmap high sampling frequency experiment, echo=FALSE, fig.width=14, fig.height=14}
fullmatrix <- correlation_matrices$correlationMatrix

# Add the dynamic range to the labels
labels <- rownames(fullmatrix)
dr_labels <- c()
for(lab in 1:length(labels)){
  dr_labels <- c(dr_labels, paste(data_high$specifications[data_high$specifications$label == labels[lab], 6], labels[lab], sep = "_"))
}
colnames(fullmatrix) <- dr_labels
rownames(fullmatrix) <- dr_labels


heatmap <- corrplot::corrplot(pmax(fullmatrix, t(fullmatrix), na.rm = TRUE), 
                                  type = 'lower', method = 'color', order = "alphabet", 
                                  number.cex = 1.25, cl.cex = 2, tl.cex = 2,
                                  col.lim = c(0,1), addCoef.col = 'black', tl.srt = 45, 
                                  diag = TRUE, tl.col = "black")
rm(fullmatrix, heatmap)
```
### Heatmap per shaker condition
```{r Heatmap high sampling frequency experiment per shaker condition, echo=FALSE, fig.width=14, fig.height=15}
fullmatrices <- correlations_perShakerCondition

# Add the dynamic range to the labels
for(matrix in 1:length(fullmatrices)){
  colnames(fullmatrices[[matrix]]$correlationMatrix) <- dr_labels
  rownames(fullmatrices[[matrix]]$correlationMatrix) <- dr_labels
}

heatmaps <- list()
for (condition in 1:length(fullmatrices)){
  heatmaps[[condition]] <- corrplot::corrplot(pmax(fullmatrices[[condition]]$correlationMatrix,
                          t(fullmatrices[[condition]]$correlationMatrix), na.rm = TRUE), 
                            title = paste0("shaker condition: ", names(fullmatrices)[condition]),
                            type = 'lower', method = 'color', order = "alphabet", 
                            number.cex = 1.25, cl.cex = 2, tl.cex = 2,
                            col.lim = c(-1,1), addCoef.col = 'black', tl.srt = 45, mar=c(0,0,2,0),
                            diag = TRUE, tl.col = "black")
}

rm(fullmatrices, heatmaps)
```

## Comparison of cross-correlation coefficients between dynamic ranges

```{r Calculate average maximum cross lagged correlation coefficients high sampling frequency experiment per shaker condition, include=FALSE}
avg_crosscor_between <- list()
avg_crosscor_within <- list()

for (condition in 1:length(correlations_perShakerCondition)){
  avg_crosscor <- dataframeShapeComparisonBeween(correlations_perShakerCondition[[condition]], data_high$specifications, between = "dynamic_range") 
  avg_crosscor_between[[condition]] <- avg_crosscor[avg_crosscor$same_group == "0",]
  avg_crosscor_within[[condition]] <- avg_crosscor[avg_crosscor$same_group == "1",]
}
rm(avg_crosscor, condition)
names(avg_crosscor_between) <- names(correlations_perShakerCondition)
names(avg_crosscor_within) <- names(correlations_perShakerCondition)
```

```{r Between dynamic range comparison high sampling frequency experiment, echo=FALSE}
table_between <- list()

for(condition in 1:length(avg_crosscor_between)){
  df_correlations <- avg_crosscor_between[[condition]]
  df_correlations$correlation <- as.numeric(df_correlations$correlation)

    
  cor24 <- rbind(df_correlations[which(df_correlations$between1 == 2 & df_correlations$between2 == 4) ,],
                 df_correlations[which(df_correlations$between1 == 4 & df_correlations$between2 == 2) ,])
  cor28 <- rbind(df_correlations[which(df_correlations$between1 == 2 & df_correlations$between2 == 8) ,],
                 df_correlations[which(df_correlations$between1 == 8 & df_correlations$between2 == 2) ,])
  cor216 <- rbind(df_correlations[which(df_correlations$between1 == 2 & df_correlations$between2 == 16) ,],
                 df_correlations[which(df_correlations$between1 == 16 & df_correlations$between2 == 2) ,])
  cor48 <- rbind(df_correlations[which(df_correlations$between1 == 4 & df_correlations$between2 == 8) ,],
                 df_correlations[which(df_correlations$between1 == 8 & df_correlations$between2 == 4) ,])
  cor416 <- rbind(df_correlations[which(df_correlations$between1 == 4 & df_correlations$between2 == 16) ,],
                 df_correlations[which(df_correlations$between1 == 16 & df_correlations$between2 == 4) ,])
  cor816 <- rbind(df_correlations[which(df_correlations$between1 == 8 & df_correlations$between2 == 16) ,],
                 df_correlations[which(df_correlations$between1 == 16 & df_correlations$between2 == 8) ,])
  
  tab <- rbind(summary(cor24$correlation), summary(cor28$correlation), summary(cor216$correlation), 
               summary(cor48$correlation), summary(cor416$correlation), summary(cor816$correlation))
  sd <- c(sd(cor24$correlation), sd(cor28$correlation), sd(cor216$correlation), 
               sd(cor48$correlation), sd(cor416$correlation), sd(cor816$correlation))
  tab <- cbind(tab, sd)
  groups <- c("2_vs_4", "2_vs_8", "2_vs_18", "4_vs_8", "4_vs_16", "8_vs_16")
  tab <- tab[,-1]
  tab <- tab[,-5]
  tab <- cbind(groups, tab)
  table_between[[condition]] <- tab
}
rm(tab, cor24, cor28, cor216, cor48, cor416, cor816, sd, groups, condition)
names(table_between) <- names(correlations_perShakerCondition)
```

```{r Print between comparison tables high frequency experiment, echo = FALSE}
for(condition in 1:length(table_between)){
  print(paste0("Shaker condition: ", names(table_between)[condition]))
  print(table_between[[condition]])
}
```

```{r Line graph high sampling frequency, echo=TRUE}
shaker_frequencies <- c(30, 40, 50, 62, 75, 87, 100, 112, 125, 137, 150, 160, 175, 187, 200, 212, 225, 237, 250)
groups <- c("2_vs_4", "2_vs_8", "2_vs_18", "4_vs_8", "4_vs_16", "8_vs_16")
comparison <- rep(groups, length(shaker_frequencies))
correlations <- c()
sd <- c()
shaker_frequency <- c()

for(condition in 1:length(table_between)){
  correlations <- c(correlations, table_between[[condition]][,4])
  sd <- c(sd, table_between[[condition]][,6])
  shaker_frequency <- c(shaker_frequency, rep(shaker_frequencies[condition], length(table_between[[condition]][,4])))
}
overview_tab_between <- cbind(comparison, correlations, shaker_frequency, sd)
overview_tab_between <- as.data.frame(overview_tab_between)

 # ggplot2::geom_errorbar(ggplot2::aes(ymin = as.numeric(correlations) - (1.96 * as.numeric(sd)),
  #                                    ymax = as.numeric(correlations) + (1.96*as.numeric(sd))), 
   #                      width=.1, position=ggplot2::position_dodge(0.5)) +

line_plot <- ggplot2::ggplot(data = overview_tab_between, ggplot2::aes(x = as.factor(shaker_frequency), y = as.numeric(correlations), group = as.factor(comparison))) +
  ggplot2::geom_line(ggplot2::aes(color = as.factor(comparison)))+
  ggplot2::geom_point(ggplot2::aes(color = as.factor(comparison))) +
  ggplot2::labs(x="Shaker frequency (rpm)", y = "Cross-correlation coefficient", color = "Dynamic range comparison") +
  ggplot2::scale_x_discrete(limits = c("30", "40", "50", "62", "75", "87", "100", "112", "125", "137", "150", "160", "175", "187", "200", "212", "225", "237", "250")) +
  ggplot2::theme_classic() +
  ggplot2::theme(legend.position="bottom") +
  ggplot2::geom_hline(yintercept=0.25, linetype="dashed") +
  ggplot2::geom_hline(yintercept=0.5, linetype="dashed") +
  ggplot2::geom_hline(yintercept=0.75, linetype="dashed")

ggplot2::ggsave(file=paste0(datadir, "plots/linegraph_E4_crosscorrelations.png"), line_plot, width = 10, height = 8, dpi = 900) #saves g
```

```{r Plot line graph, echo=TRUE}
plot(line_plot)
```

