---
title: "Frequency domain analyses"
author: "Annelinde Lettink"
output: 
  html_notebook:
    toc: true
    toc_depth: 6  # upto four depths of headings (specified by #, ##, ###, ####, ##### and ######)
---

```{r Setup and load data, include=FALSE}
#install.packages("knitr")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'pdf')

rm(list=ls())
graphics.off()

# Update to be your local directory, if all goes well this is the only line you will have to update
shaker_experiments_folder = "/Users/annelindelettink/Documents/Work MacBook Pro Annelinde/Mechanical Shaker Machine"
# shaker_experiments_folder = "~/data/VUMC/shaker_experiments"

#====================================================================================
# Specify file paths
datadir = paste0(shaker_experiments_folder, "/analyses/")
if (!dir.exists(datadir)) {
  stop(paste0("Directory does not exist: ", datadir))
}

# Load cleaned mechanical shaker machine data for both experiments
filename_high = paste0(datadir, "ms_flat_HA_cleaned_high.RData") 
filename_low = paste0(datadir, "ms_flat_HA_cleaned_low.RData") 
if (!file.exists(filename_high)) {
  stop(paste0("File does not exist: ", filename_high))
}
load(filename_high)
if (!file.exists(filename_low)) {
  stop(paste0("File does not exist: ", filename_low))
}
load(filename_low)
rm(filename_high, filename_low)
```

```{r Functions, include=FALSE}
#' deriveSpectrum
#'
#' @description 'deriveSpectrum' derives the frequency spectrum of a signal
#'
#' @param x Vector that contains the values of the signal (i.e., acceleration)
#' @param sampling_frequency Integer that indicates the sampling frequency of the signal (i.e., the average number of samples obtained in one second in Hz)
#' @param file_name String to indicate the name of the plot (i.e., "spectrum" result in the following name for plot 1 "spectrum_1.jpeg").
#' @param plot Boolean, default is TRUE the derived spectrum will be plotted
#' @param XLIM Vector that limits the frequency content of the plot, default is c(0, 5) to limit the frequency content to 5 Hz (as 250 rpm / 60 = 4.1667 Hz is the expected max)
#' @return An object of class "spec", which is a list containing at least the following components: 
#' \item{freq}{Vector of frequencies at which the spectral density is estimated.} 
#' \item{spec}{Vector (for univariate series) or matrix (for multivariate series) of estimates of the spectral density at frequencies corresponding to freq.}
#' \item{coh}{NULL for univariate series. For multivariate time series, a matrix containing the squared coherency between different series.}  
#' \item{phase}{NULL for univariate series. For multivariate time series a matrix containing the cross-spectrum phase between different series.} 
#' \item{series}{The name of the time series.}
#' \item{snames}{For multivariate input, the names of the component series.}
#' \item{method}{The method used to calculate the spectrum.} 
#' @export
deriveSpectrum <- function(x, sampling_frequency, file_name, plot = TRUE, XLIM = c(0, 5)) {
  specd <- spectrum(x, log = "no", plot = FALSE) # Default calculate spectrum on log-scale, but we use raw data
  
  # Frequency axis of the spectrum, specd, is calculated in terms of cycles per sampling interval
  delta <- 1/sampling_frequency
  specd$specx <- specd$freq/delta # Converts to cycles per unit time (divide by sampling interval)
  specd$specy <- 2*specd$spec # Multiply the spectrum by 2 to make it actually equal to variance
  
  #Save the frequency spectrum plot and zoom in on low-frequencies (at most 5 based on max rpm)
  if (plot == TRUE){
    jpeg(paste(datadir, paste0(paste(file_name, file, sep = "_"), ".jpeg"), sep = "/plots/"),
         width=600, height=500, res=120) # start export
    plot(specd$specx, specd$specy,
         main = "HA", xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l") 
    dev.off()
  }
  return(specd)
}

#' deriveMeanSpectrum
#'
#' @description 'deriveMeanSpectrum' derives the mean frequency spectrum from the derived spectra
#'
#' @param spectra A list containing the raw spectra over which the mean will be derived
#' @return A list containing two following elements: 
#' \item{x}{Vector of frequencies at which the spectral density (y) is estimated} 
#' \item{y}{Vector of estimates of the spectral density at frequencies corresponding to frequencies (x)}
#' @export
deriveMeanSpectrum <- function(spectra) {
  df_specx <- data.frame()
  df_specy <- data.frame()
  for(spectrum in 1:length(spectra)){
    df_specx <- rbind(df_specx, spectra[[spectrum]]$specx)
    df_specy <- rbind(df_specy, spectra[[spectrum]]$specy)
  }
  x <- colMeans(df_specx)
  y <- colMeans(df_specy)
  spec <- list(x = x, y = y)
  return(spec)
}

#' derivePeaks
#'
#' @description 'derivePeaks' derives peaks from the mean frequency spectrum
#'
#' @param mean.spec A list containing estimates of the spectral density (y) at corresponding frequencies (x)
#' @param sampling_frequency An integer indicating whether peaks are derived for the low or high sampling frequency experiment, one of c(25, 100) 
#' @param XLIM Vector that limits the frequency content, default is c(0, 5) to limit the frequency content to 5 Hz (as 250 rpm / 60 = 4.1667 Hz is the expected max)
#' @param plot Boolean, default is TRUE the derived peaks will be plotted over the mean frequency spectrum
#' @return A vector indicating the frequencies at which the peaks occur 
#' @export
derivePeaks <- function(mean.spec, sampling_frequency, XLIM = c(0, 5), plot = TRUE) {
  peaks <- pracma::findpeaks(mean.spec$y[mean.spec$x <= XLIM[2]], 
                             minpeakheight = mean(mean.spec$y) + sd(mean.spec$y) , 
                             minpeakdistance = 4*sampling_frequency)
  local.max <- mean.spec$x[sort(peaks[,2])]
  local.max <- local.max[local.max <= XLIM[2]]
  
  if (plot == TRUE){
    plot <- plot(mean.spec$x, mean.spec$y,  xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l", main = "Detected peaks")
    for(max in 1:length(local.max)){
      p <- plot + abline(v=local.max[max], col = "red")
    }
  }
  return(local.max)
}

#' deriveBins
#'
#' @description 'deriveBins' derives frequency bins from the peaks detected in the mean frequency spectrum
#'
#' @param peaks A vector containing the corresponding frequencies of where the peaks (local maxima) are located
#' @param mean.spec A list containing estimates of the spectral density (y) at corresponding frequencies (x)
#' @param XLIM Vector that limits the frequency content, default is c(0, 5) to limit the frequency content to 5 Hz (as 250 rpm / 60 = 4.1667 Hz is the expected max)
#' @param plot Boolean, default is TRUE the derived bins will be plotted over the mean frequency spectrum
#' @return A vector indicating the frequencies at which the peaks occur 
#' @export
deriveBins <- function(peaks, mean.spec, XLIM = c(0, 5), plot = TRUE) {
  peaks <- c(XLIM[1], peaks)
  
  freq.bin <- c()
  for(bin in 2:length(peaks)){
    freq.bin <- c(freq.bin, (peaks[bin-1] + peaks[bin])/2)
  }
  if(plot == TRUE){
    plot <- plot(mean.spec$x, mean.spec$y,  xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l", main = "Derived frequency cut-offs")
    for(bin in 1:length(freq.bin)){
      p <- plot + abline(v=freq.bin[bin], col = "red")
    }  
  }
  freq.bin <- c(XLIM[1], freq.bin, XLIM[2])
  return(freq.bin)
}

#' deriveComparisonValues
#'
#' @description 'deriveComparisonValues' derives the dominant frequency and the mean power spectral density from a frequency spectrum
#'
#' @param spectrum An object of class "spec"
#' @param freqBins Vector containing frequency values that define the frequency bins (can be derived using the function 'deriveBins') 
#' @return Data.frame consisting of two columns: \item{domFreq}{The dominant frequency of the signal in each frequency bin} \item{meanPSD}{The mean power spectral density of the signal in each frequency bin}
#' @export
deriveComparisonValues <- function(spectrum, freqBins){
  df <- data.frame()
  for(i in 1:(length(freqBins)-1)){
    lower <- freqBins[i]
    upper <- freqBins[i+1]
    index <- which(spectrum$specx >= lower & spectrum$specx <= upper)
    df[i, 1] <- spectrum$specx[index[which.max(spectrum$specy[index])]] # dominant frequency
    df[i, 2] <- mean(spectrum$specy[index]) # mean power spectral density
  }
  colnames(df) <- c("domFreq", "meanPSD")
  return(df)
}

#' createBoxplot
#'
#' @description 'createBoxplot' creates a boxplot between or within devices for the dominant frequency or the mean power spectral density per frequency bin
#'
#' @param data A data set containing the comparison values for the experiment
#' @param freqBins A vector that indicates the cut-offs for the freqency bins
#' @param outcome String containing the type of the comparison value. One of c("domFreq", "meanPSD") 
#' @param experiment String to indicate the experiment. One of c("low", "high") 
#' @param relative Boolean to indicate whether the values need to be plotted relatively to the frequency bin mean. One of c("TRUE", "FALSE") 
#' @return Boxplot between or within devices for the outcome (y-axis) per frequency bin (x-axis)
#' @export
createBoxplot <- function(data, freqBins, outcome = c("domFreq", "meanPSD"), experiment = c("low", "high"), relative = c(TRUE, FALSE)){
  if (outcome == "domFreq" & relative == FALSE){
    label = "Dominant frequency (Hertz)"
  } else if (outcome == "meanPSD" & relative == FALSE){
    label = "Mean power spectral density (g2/Hertz)"
  } else if (outcome == "domFreq" & relative == TRUE){
    label = "Deviation from mean bin dominant frequency (%)"
  } else {
    label = "Deviation from mean bin power spectral density (%)"
  }
  if (experiment == "low"){
    bins = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
      } else if (experiment == "high"){
    bins = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13")
    }
  bin.labels <- c()
  for (bin in 2:length(freqBins)){
    bin.labels <- c(bin.labels, paste(round(freqBins[bin-1], digits = 2), 
                                      round(freqBins[bin], digits = 2), sep = "-"))
  }
  if(relative == TRUE){
    bin.means <- aggregate(data[, c("domFreq", "meanPSD")], list(data$bin), mean)
    if(outcome == "domFreq"){
        data <- transform(data, domFreq.binmean = ave(domFreq, bin.labels, FUN = mean))
        data$rel <- (abs(data$domFreq - data$domFreq.binmean)/data$domFreq.binmean)*100
    } else{
        data <- transform(data, meanPSD.binmean = ave(meanPSD, bin.labels, FUN = mean))
        data$rel <- (abs(data$meanPSD - data$meanPSD.binmean)/data$meanPSD.binmean)*100
      #data <- transform(data, perc = ave(meanPSD, bin.labels, FUN = prop.table)*100)
    }
    outcome = "rel"
  }
  
  boxplot <- ggpubr::ggboxplot(data, x = "bin", y = outcome, color = "brand",
                               xlab = "Frequency bin (Hertz)", ylab = label, 
                                order = bins,
                               font.label = list(size = 8, color = "black")) +
    ggpubr::font("xlab", size = 6, color = "black") +
    ggpubr::font("x.text", size = 8, color = "black") + 
    ggpubr::font("ylab", size = 6, color = "black") +
    ggpubr::font("y.text", size = 8, color = "black") + 
    ggpubr::font("legend.title", size = 8, color = "black") +
    ggpubr::font("legend.text", size = 8, color = "black") + 
    ggplot2::scale_x_discrete(breaks=bins,
                     labels= bin.labels)
  return(boxplot)
}

#' evaluateMM
#'
#' @description 'evaluateMM' function to indicate if the likelihood ratio test was significant
#'
#' @param model1 Model object with the least parameters
#' @param model2 Model object with the most parameters
#' @param type String containing the addition to the model. One of c("randomIntercept", "randomSlope") 
#' @return Boolean: \item{TRUE}{The addition of a random intercept/slope is necessary} \item{FALSE}{The addition of a random intercept/slope is not necessary}
#' @export
evaluateMM <- function(model1, model2, type = c("randomIntercept", "randomSlope")){
  logLikDiff <- (-2*(logLik(model1))) - (-2*logLik(model2))
  if(type == "randomIntercept"){
    return(logLikDiff > 3.84)
  }
  else{ 
    return(logLikDiff > 5.99)
  }
}

#' getOutputMM
#'
#' @description 'getOutputMM' function print the model output
#'
#' @param finalModel Model object for which the output will be printed
#' @return Printed model output including fixed and random effect estimates
#' @export
getOutputMM <- function(finalModel) {
  print(summary(finalModel)) #fixed effect estimates
  lme4::VarCorr(finalModel) #random effect estimates
}
```

# Frequency spectra

This section presents frequency spectra derived from the raw acceleration signals for both experiments. We used the mean frequency spectra to the derive frequency bins for both experiments, over which aggregated outcome values (i.e., dominant frequency and mean power spectral density) were calculated for statistical comparison between brands.

```{r Settings for frequency spectra, include=FALSE}
# Plot parameters
plot = TRUE
xlabel = "Frequency (Hertz)"
ylabel = "Spectral Density (g2/Hertz)"
XLIM = c(0, 5) # Limit frequency content to 5 Hz (as 250 rpm / 60 = 4.1667 Hz is the expected max)
```

## Mean frequency spectra
This section presents the mean frequency spectra for both experiments separately.
### Low sampling frequency experiment

```{r Derive frequency spectra low sampling frequency experiment, echo=FALSE}
# DERIVE RAW FREQUENCY SPECTRA (for visual inspection and save data for further analysis)
raw.spectra_low <- list() 

for (file in 1:length(clean_data_low$data)) { # for all accelerometer files
  sampling_frequency <- as.numeric(clean_data_low$specifications$sampling_frequency[file])  
  HA <- clean_data_low$data[[file]]$HA #for horizontal axis
  rawSpectrum <- deriveSpectrum(HA, sampling_frequency, file_name = names(clean_data_low$data)[[file]])
  raw.spectra_low[[file]] <- rawSpectrum
}
rm(rawSpectrum, file)
names(raw.spectra_low) <- clean_data_low$specifications$label  

mean.freqspec_low <- deriveMeanSpectrum(raw.spectra_low)
```

```{r Plot mean frequency spectrum low sampling frequency, echo=FALSE}
plot(mean.freqspec_low$x, mean.freqspec_low$y, xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l", main = "Mean frequency spectrum low sampling frequency experiment")
```

### High sampling frequency experiment

```{r Derive frequency spectra high sampling frequency experiment, echo=FALSE}
# DERIVE RAW FREQUENCY SPECTRA (for visual inspection and save data for further analysis)
raw.spectra_high <- list() 

for (file in 1:length(clean_data_high$data)) { # for all accelerometer files
  sampling_frequency <- as.numeric(clean_data_high$specifications$sampling_frequency[file])  
  HA <- clean_data_high$data[[file]]$HA #for horizontal axis
  rawSpectrum <- deriveSpectrum(HA, sampling_frequency, file_name = names(clean_data_high$data)[[file]])
  raw.spectra_high[[file]] <- rawSpectrum
}
rm(rawSpectrum, file)
names(raw.spectra_high) <- clean_data_high$specifications$label

mean.freqspec_high <- deriveMeanSpectrum(raw.spectra_high)
```

```{r Plot mean frequency spectrum high sampling frequency, echo=FALSE}
plot(mean.freqspec_high$x, mean.freqspec_high$y, xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l", main = "Mean frequency spectrum high sampling frequency experiment")
```

```{r Save raw spectra, include=FALSE}
raw_spectra <- list(high = raw.spectra_high, high_mean = mean.freqspec_high, specifications_high = clean_data_high$specifications,
                    low = raw.spectra_low, low_mean = mean.freqspec_low, specifications_low = clean_data_low$specifications) # Save derived raw spectra for HA in one list
save(raw_spectra, file = paste0(datadir, "frequency_spectra.RData"))
rm(raw.spectra_high, raw.spectra_low, HA, sampling_frequency, mean.freqspec_high, mean.freqspec_low)
```

## Frequency bins
We detected the frequency peaks in the mean frequency spectra both experiments separately, and derived the frequencies in Hertz associated with these peaks. We divided these frequencies by two to derive the frequency cut-offs for the bins to ensure the peaks fell within these bins. Aggregated outcome values (i.e., dominant frequency and mean power spectral density) were calculated over the frequency bins for statistical comparison between brands.

### Derive frequency bins

#### Low sampling frequency experiment
```{r Derive frequency bins low sampling frequency, echo=FALSE}
peaks_low <- derivePeaks(raw_spectra$low_mean, 25, XLIM, plot = TRUE)
bins_low <- deriveBins(peaks_low, raw_spectra$low_mean, XLIM = c(0, 5), plot = TRUE) 
```
#### High sampling frequency experiment
```{r Derive frequency bins high sampling frequency, echo=FALSE}
peaks_high <- derivePeaks(raw_spectra$high_mean, 100, XLIM, plot = TRUE)
bins_high <- deriveBins(peaks_high, raw_spectra$high_mean, XLIM = c(0, 5), plot = TRUE) 
```
### Frequency spectrum plots
This section presents the frequency spectrum plots of the overlayed device acceleration signals and for all signals separately, where the frequency bins are indicated by red vertical lines.

#### Low sampling frequency experiment
```{r Overlay all spectra low sampling frequency experiment, echo=FALSE}
plot <- plot(raw_spectra$low[[1]]$specx, raw_spectra$low[[1]]$specy, main = "Overlay all frequency spectra", 
       xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l")
for(i in 1:length(bins_low)){
    plot <- plot + abline(v=bins_low[i], col = "red")
  }  

for (spec in 2:length(raw_spectra$low)) {
  plot + lines(raw_spectra$low[[spec]]$specx, raw_spectra$low[[spec]]$specy, type = "l")
  
}
rm(spec, i)
```

```{r Plot spectra low sampling frequency experiment, echo=FALSE}
for (spec in 1:length(raw_spectra$low)) {
  spectrum <- raw_spectra$low[[spec]]
  title <- paste("Frequency spectrum plot: ", raw_spectra$specifications_low$label[[spec]])
  plot(spectrum$specx, spectrum$specy, main = title, 
       xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l")
  for(i in 1:length(bins_low)){
    abline(v=bins_low[i], col = "red")
  }  
}
rm(spec, title, spectrum, i)
```

#### High sampling frequency experiment
```{r Overlay all spectra high sampling frequency experiment, echo=FALSE}
plot <- plot(raw_spectra$high[[1]]$specx, raw_spectra$high[[1]]$specy, main = "Overlay all frequency spectra", 
       xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l")
for(i in 1:length(bins_high)){
    plot <- plot + abline(v=bins_high[i], col = "red")
  }  

for (spec in 2:length(raw_spectra$high)) {
  plot + lines(raw_spectra$high[[spec]]$specx, raw_spectra$high[[spec]]$specy, type = "l")
  
}
rm(spec, i)
```

```{r Plot spectra high sampling frequency experiment, echo=FALSE}
#load(paste0(datadir, "frequency_spectra.RData"))

for (spec in 1:length(raw_spectra$high)) {
  title <- paste("Frequency spectrum plot: ", raw_spectra$specifications_high$label[[spec]])
  plot(raw_spectra$high[[spec]]$specx, raw_spectra$high[[spec]]$specy, main = title, 
       xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l") 
  for(i in 1:length(bins_high)){
    abline(v=bins_high[i], col = "red")
  }  
}
rm(spec, i)
```

```{r Compute dominant frequency (domFreq) and mean power spectral density (meanPSD) for HA , include = FALSE}
comparison_values_low <- list()
for(spectrum in 1:length(raw_spectra$low)){ 
  comparison_values_low[[spectrum]] <- deriveComparisonValues(raw_spectra$low[[spectrum]], bins_low)
}
names(comparison_values_low) <- names(raw_spectra$low)

comparison_values_high <- list()
for(spectrum in 1:length(raw_spectra$high)){ 
  comparison_values_high[[spectrum]] <- deriveComparisonValues(raw_spectra$high[[spectrum]], bins_high)
}
names(comparison_values_high) <- names(raw_spectra$high)
#cat("\nSaving data...")
comparison_values <- list(high = comparison_values_high, low = comparison_values_low)
save(comparison_values, file = paste0(datadir, "domFreq_meanPSD.RData")) 

rm(spectrum, comparison_values_high, comparison_values_low)
```

```{r WIDE FORMAT: Prepare data set for statistical analyses, include = FALSE}
# Wide format: for both experiments and outcomes domFreq and meanPSD

df_wide_low <- data.frame()
id <- names(comparison_values$low)
df_wide_low <- cbind(id, clean_data_low$specifications$brand, clean_data_low$specifications$sampling_frequency,
                 clean_data_low$specifications$dynamic_range)
domFreq_wide_low <- data.frame() 
meanPSD_wide_low <- data.frame()
for (accelerometer in 1:length(comparison_values$low)) {
  domFreq_wide_low <- rbind(domFreq_wide_low, comparison_values$low[[accelerometer]]$domFreq)
  meanPSD_wide_low <- rbind(meanPSD_wide_low, comparison_values$low[[accelerometer]]$meanPSD)
}
domFreq_ms_wide_low <- cbind(df_wide_low, domFreq_wide_low)
colnames(domFreq_ms_wide_low) <- c("serialnumber", "brand", "sampling_frequency", "dynamic_range", 
                               "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
meanPSD_ms_wide_low <- cbind(df_wide_low, meanPSD_wide_low)
colnames(meanPSD_ms_wide_low) <- c("serialnumber", "brand", "sampling_frequency", "dynamic_range", 
                               "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
rm(df_wide_low, id, domFreq_wide_low, meanPSD_wide_low, accelerometer)

df_wide_high <- data.frame()
id <- names(comparison_values$high)
df_wide_high <- cbind(id, clean_data_high$specifications$brand, clean_data_high$specifications$sampling_frequency,
                 clean_data_high$specifications$dynamic_range)
domFreq_wide_high <- data.frame() 
meanPSD_wide_high <- data.frame()
for (accelerometer in 1:length(comparison_values$high)) {
  domFreq_wide_high <- rbind(domFreq_wide_high, comparison_values$high[[accelerometer]]$domFreq)
  meanPSD_wide_high <- rbind(meanPSD_wide_high, comparison_values$high[[accelerometer]]$meanPSD)
}
domFreq_ms_wide_high <- cbind(df_wide_high, domFreq_wide_high)
colnames(domFreq_ms_wide_high) <- c("serialnumber", "brand", "sampling_frequency", "dynamic_range", 
                               "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13")
meanPSD_ms_wide_high <- cbind(df_wide_high, meanPSD_wide_high)
colnames(meanPSD_ms_wide_high) <- c("serialnumber", "brand", "sampling_frequency", "dynamic_range", 
                               "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13")
rm(df_wide_high, id, domFreq_wide_high, meanPSD_wide_high, accelerometer, comparison_values, clean_data_high, clean_data_low)
```

```{r LONG FORMAT: Prepare data set for statistical analyses, include = FALSE}
# Combining domFreq and meanPSD (gather columns for outcome variables)
library(tidyverse)

domFreq_long_low <- domFreq_ms_wide_low %>%
  gather(key = "bin", value = "domFreq", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12") %>%
  rstatix::convert_as_factor(serialnumber, bin, brand, sampling_frequency, dynamic_range)
meanPSD_long_low <- meanPSD_ms_wide_low %>%
  gather(key = "bin", value = "meanPSD", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12") %>%
  rstatix::convert_as_factor(serialnumber, bin, brand, sampling_frequency, dynamic_range)
df_low <- merge(domFreq_long_low, meanPSD_long_low) # Join both data.frame objects to derive a single dataset
rm(domFreq_long_low, domFreq_ms_wide_low, meanPSD_long_low, meanPSD_ms_wide_low)
df_low <- df_low[order(df_low$serialnumber),]

domFreq_long_high <- domFreq_ms_wide_high %>%
  gather(key = "bin", value = "domFreq", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13") %>%
  rstatix::convert_as_factor(serialnumber, bin, brand, sampling_frequency, dynamic_range)
meanPSD_long_high <- meanPSD_ms_wide_high %>%
  gather(key = "bin", value = "meanPSD", "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13") %>%
  rstatix::convert_as_factor(serialnumber, bin, brand, sampling_frequency, dynamic_range)
df_high <- merge(domFreq_long_high, meanPSD_long_high) # Join both data.frame objects to derive a single dataset
rm(domFreq_long_high, domFreq_ms_wide_high, meanPSD_long_high, meanPSD_ms_wide_high)
df_high <- df_high[order(df_high$serialnumber),]

df_long <- list(low = df_low, high = df_high)
#cat("\nSaving data...")
save(df_long, file = paste0(datadir, "domfreq_meanPSD_long.RData"))
rm(df_high, df_low)
```

# Statistical comparison

This section presents the statistical comparison of the frequency spectra in terms of the aggregated dominant frequency and mean power spectral density for each frequency bin per sampling frequency experiment.

## Low sampling frequency experiment

### Boxplots
This section presents the boxplots for both outcomes (dominant frequency and mean power spectral density) for the low sampling frequency experiment.

```{r Create boxplots low sampling frequency experiment, include=FALSE}
#load(paste0(datadir, "domfreq_meanPSD_long.RData"))

bxp_domfreq_low <- createBoxplot(df_long$low, bins_low, outcome = "domFreq", experiment = "low", relative = FALSE)
bxp_psd_low <- createBoxplot(df_long$low, bins_low, outcome = "meanPSD", experiment = "low", relative = FALSE)

gridExtra::grid.arrange(bxp_domfreq_low, bxp_psd_low, nrow=1) #arranges plots within grid
g <- gridExtra::arrangeGrob(bxp_domfreq_low + ggplot2::theme(legend.position="none"), 
                            bxp_psd_low + ggplot2::theme(legend.position="bottom"), nrow=2) #generates g 
ggplot2::ggsave(file=paste0(datadir, "plots/boxplots_HA_LOW.pdf"), g) #saves g

bxp_domfreq_low_perc <- createBoxplot(df_long$low, bins_low, outcome = "domFreq", experiment = "low", relative = TRUE)
bxp_psd_low_perc <- createBoxplot(df_long$low, bins_low, outcome = "meanPSD", experiment = "low", relative = TRUE)

gridExtra::grid.arrange(bxp_domfreq_low_perc, bxp_psd_low_perc, nrow=1) #arranges plots within grid
g_perc <- gridExtra::arrangeGrob(bxp_domfreq_low_perc + ggplot2::theme(legend.position="none"),
                                bxp_psd_low_perc + ggplot2::theme(legend.position="bottom"),
                                nrow=2)#generates g_perc
ggplot2::ggsave(file=paste0(datadir, "plots/boxplots_HA_LOW_percentage.pdf"), g_perc) #saves g_perc
```

Boxplot
```{r Print boxplot low sampling frequency experiment, echo=FALSE}
plot(g)
```

Boxplot with percentage deviation from bin means
```{r Print percentage boxplot low sampling frequency experiment, echo=FALSE}
plot(g_perc)
```

### Mixed model analyses
This section presents the statistical comparison between the brand for the low sampling frequency experiment. Thee mixed models for mean power spectral density and dominant frequency with a random intercept for serial number are presented separately.

```{r Mixed Model Analyses low sampling frequency experiment, include=FALSE}
# Starting model: with random intercept for accelerometer id (to adjust for the repeated measurements of the accelerometer devices)
meanPSD_intercept_id_low <- nlme::lme(meanPSD ~ brand, random = ~ 1|serialnumber, 
                                      method="ML", data = df_long$low)
domFreq_intercept_id_low <- nlme::lme(domFreq ~ brand, random = ~ 1|serialnumber, 
                                      method="ML", data = df_long$low)

# Add random intercept for frequency bin (to adjust for the frequencies)
meanPSD_intercept_bin_low <- nlme::lme(meanPSD ~ brand, random = list(~ 1|serialnumber, ~ 1|bin),
                                           method="ML", data = df_long$low)
domFreq_intercept_bin_low <- nlme::lme(domFreq ~ brand, random = list(~ 1|serialnumber, ~ 1|bin),
                                           method="ML", data = df_long$low)

# Evaluate if adding a random intercept for frequency bin is necessary
evaluateMM(meanPSD_intercept_id_low, meanPSD_intercept_bin_low, type = "randomIntercept") #FALSE adding a random intercept for frequency bin to the model is not necessary
evaluateMM(domFreq_intercept_id_low, domFreq_intercept_bin_low, type = "randomIntercept") #FALSE adding a random intercept for frequency bin to the model is not necessary
rm(meanPSD_intercept_bin_low, domFreq_intercept_bin_low)

# Do brands differ in dominant frequency or density depending on the movement frequency (add interaction between brand and bin)
meanPSD_intercept_id_interaction_low <- nlme::lme(meanPSD ~ brand + bin + brand * bin, random = ~ 1|serialnumber, 
                                      method="ML", data = df_long$low)
domFreq_intercept_id_interaction_low <- nlme::lme(domFreq ~ brand + bin + brand * bin, random = ~ 1|serialnumber, 
                                      method="ML", data = df_long$low)
```

#### Model assumption: normal distribution of residuals

```{r Mean PSD Model assumptions low sampling frequency experiment, echo=FALSE}
ggpubr::ggqqplot(df_long$low, "meanPSD", facet.by = "bin", title = "QQ-plot mean power spectral density")
residuals_meanPSD <- resid(meanPSD_intercept_id_low)
#summary(residuals_meanPSD)
hist(residuals_meanPSD)
```

```{r Dominant frequency Model assumptions low sampling frequency experiment, echo=FALSE}
ggpubr::ggqqplot(df_long$low, "domFreq", facet.by = "bin", title = "QQ-plot dominant frequency")
residuals_domFreq <- resid(domFreq_intercept_id_low)
#summary(residuals_domFreq)
hist(residuals_domFreq) # approximately normally distributed
```

#### Results Mean Power Spectral Density

##### Model: meanPSD ~ brand
###### Reference: ActiGraph 
```{r Report results for Mean PSD low sampling frequency experiment: ActiGraph, echo=FALSE}
getOutputMM(meanPSD_intercept_id_low)
```

###### Reference: MOX
```{r Report results for Mean PSD low sampling frequency experiment: MOX, echo=FALSE}
df_long$low$brand <- relevel(df_long$low$brand, ref = 5) 
meanPSD_intercept_id_low <- nlme::lme(meanPSD ~ brand, random = ~ 1|serialnumber, 
                                      method="ML", data = df_long$low)
getOutputMM(meanPSD_intercept_id_low)
```
###### Reference: GENEActiv
```{r Report results for Mean PSD low sampling frequency experiment: GENEActiv, echo=FALSE}
df_long$low$brand <- relevel(df_long$low$brand, ref = 5) 
meanPSD_intercept_id_low <- nlme::lme(meanPSD ~ brand, random = ~ 1|serialnumber, 
                                      method="ML", data = df_long$low)
getOutputMM(meanPSD_intercept_id_low)
```
###### Reference: Axivity
```{r Report results for Mean PSD low sampling frequency experiment: Axivity, echo=FALSE}
df_long$low$brand <- relevel(df_long$low$brand, ref = 5) 
meanPSD_intercept_id_low <- nlme::lme(meanPSD ~ brand, random = ~ 1|serialnumber, 
                                      method="ML", data = df_long$low)
getOutputMM(meanPSD_intercept_id_low)
```
###### Reference: activPAL
```{r Report results for Mean PSD low sampling frequency experiment: activPAL, echo=FALSE}
df_long$low$brand <- relevel(df_long$low$brand, ref = 5) 
meanPSD_intercept_id_low <- nlme::lme(meanPSD ~ brand, random = ~ 1|serialnumber, 
                                      method="ML", data = df_long$low)
getOutputMM(meanPSD_intercept_id_low)
```

##### Model: meanPSD ~ brand + bin + brand*bin
###### Reference: ActiGraph, bin1
```{r Report results for Mean PSD interaction low sampling frequency experiment: ActiGraph, echo=FALSE}
getOutputMM(meanPSD_intercept_id_interaction_low)
```

###### Reference: MOX
```{r Report results for Mean PSD interaction low sampling frequency experiment: ActiGraph, echo=FALSE}
df_long$low$bin <- relevel(df_long$low$bin, ref = 2) 
meanPSD_intercept_id_interaction_low <- nlme::lme(meanPSD ~ brand + bin + brand*bin, random = ~ 1|serialnumber, 
                                      method="ML", data = df_long$low)
getOutputMM(meanPSD_intercept_id_interaction_low)
```

#### Results Dominant Frequency
##### Model: domFreq ~ brand
###### Reference: ActiGraph
```{r Report results for Dominant frequency low sampling frequency experiment: ActiGraph, echo=FALSE}
df_long$low$brand <- relevel(df_long$low$brand, ref = 5) 
domFreq_intercept_id_low <- nlme::lme(domFreq ~ brand, random = ~ 1|serialnumber, 
                                      method="ML", data = df_long$low)
getOutputMM(domFreq_intercept_id_low)
```
###### Reference: MOX
```{r Report results for Dominant frequency low sampling frequency experiment: MOX, echo=FALSE}
df_long$low$brand <- relevel(df_long$low$brand, ref = 5) 
domFreq_intercept_id_low <- nlme::lme(domFreq ~ brand, random = ~ 1|serialnumber, 
                                      method="ML", data = df_long$low)
getOutputMM(domFreq_intercept_id_low)
```
###### Reference: GENEActiv
```{r Report results for Dominant frequency low sampling frequency experiment: GENEActiv, echo=FALSE}
df_long$low$brand <- relevel(df_long$low$brand, ref = 5) 
domFreq_intercept_id_low <- nlme::lme(domFreq ~ brand, random = ~ 1|serialnumber, 
                                      method="ML", data = df_long$low)
getOutputMM(domFreq_intercept_id_low)
```
###### Reference: Axivity

```{r Report results for Dominant frequency low sampling frequency experiment: Axivity, echo=FALSE}
df_long$low$brand <- relevel(df_long$low$brand, ref = 5) 
domFreq_intercept_id_low <- nlme::lme(domFreq ~ brand, random = ~ 1|serialnumber, 
                                      method="ML", data = df_long$low)
getOutputMM(domFreq_intercept_id_low)
```
###### Reference: activPAL

```{r Report results for Dominant frequency low sampling frequency experiment: activPAL, echo=FALSE}
df_long$low$brand <- relevel(df_long$low$brand, ref = 5) 
domFreq_intercept_id_low <- nlme::lme(domFreq ~ brand, random = ~ 1|serialnumber, 
                                      method="ML", data = df_long$low)
getOutputMM(domFreq_intercept_id_low)
```
##### Model: domFreq ~ brand + bin + brand*bin
###### Reference: ActiGraph, bin1
```{r Report results for domintant frequency interaction low sampling frequency experiment: MOX, echo=FALSE}
getOutputMM(domFreq_intercept_id_interaction_low)
```

## High sampling frequency experiment

### Boxplots
This section presents the boxplots for both outcomes (dominant frequency and mean power spectral density) for the high sampling frequency experiment.

```{r Create boxplots high sampling frequency, include=FALSE}
#load(paste0(datadir, "domfreq_meanPSD_ms_HA_dataset_long_hf.RData"))
bxp_domfreq_high <- createBoxplot(df_long$high, bins_high, outcome = "domFreq", experiment = "high", relative = FALSE)
bxp_psd_high <- createBoxplot(df_long$high, bins_high, outcome = "meanPSD", experiment = "high", relative = FALSE) 

gridExtra::grid.arrange(bxp_domfreq_high, bxp_psd_high, nrow=1) #arranges plots within grid
g <- gridExtra::arrangeGrob(bxp_domfreq_high + ggplot2::theme(legend.position="none"), 
                            bxp_psd_high + ggplot2::theme(legend.position="bottom"), nrow=2) #generates g
ggplot2::ggsave(file=paste0(datadir, "plots/boxplots_HA_HIGH.pdf"), g) #saves g

bxp_domfreq_high_perc <- createBoxplot(df_long$high, bins_high, outcome = "domFreq", experiment = "high", relative = TRUE)
bxp_psd_high_perc <- createBoxplot(df_long$high, bins_high, outcome = "meanPSD", experiment = "high", relative = TRUE)
gridExtra::grid.arrange(bxp_domfreq_high_perc, bxp_psd_high_perc, 
                        nrow=1) #arranges plots within grid
g_perc <- gridExtra::arrangeGrob(bxp_domfreq_high_perc + ggplot2::theme(legend.position="none"),
                                bxp_psd_high_perc + ggplot2::theme(legend.position="bottom"), 
                                nrow=2) #generates g_perc
ggplot2::ggsave(file=paste0(datadir, "plots/boxplots_HA_HIGH_percentage.pdf"), g_perc) #saves g_perc
```

Boxplot

```{r Print boxplot high sampling frequency experiment, echo=FALSE}
plot(g)
```

Boxplot deviating percentage from bin means

```{r Print log boxplot high sampling frequency experiment, echo=FALSE}
plot(g_perc)
```

### Mixed model analyses
This section presents the statistical comparison between the brand for the high sampling frequency experiment. Thee mixed models for mean power spectral density and dominant frequency with a random intercept for serial number (id) are presented separately.

```{r Mixed Model Analyses high sampling frequency experiment, include=FALSE}
# Starting model: with random intercept for accelerometer id (to adjust for the repeated measurements of the accelerometer devices)
meanPSD_intercept_id_high <- nlme::lme(meanPSD ~ brand, random = ~ 1|serialnumber, 
                                       method="ML", data = df_long$high)
domFreq_intercept_id_high <- nlme::lme(domFreq ~ brand, random = ~ 1|serialnumber, 
                                       method="ML", data = df_long$high)

# Add random intercept for bin (to adjust for the frequencies)
meanPSD_intercept_bin_high <- nlme::lme(meanPSD ~ brand, random = list(~ 1|serialnumber, ~ 1|bin), 
                                            method="ML", data = df_long$high)
domFreq_intercept_bin_high <- nlme::lme(domFreq ~ brand, random = list(~ 1|serialnumber, ~ 1|bin), 
                                            method="ML", data = df_long$high)

# Evaluate if adding a random intercept for bin is necessary
evaluateMM(meanPSD_intercept_id_high, meanPSD_intercept_bin_high, type = "randomIntercept") #FALSE adding a random intercept for frequency bin to the model is not necessary
evaluateMM(domFreq_intercept_id_high, domFreq_intercept_bin_high, type = "randomIntercept") #FALSE adding a random intercept for frequency bin to the model is not necessary
rm(meanPSD_intercept_bin_high, domFreq_intercept_bin_high)

# Do brands differ in dominant frequency or density depending on the movement frequency (add interaction between brand and bin)
meanPSD_intercept_id_interaction_high <- nlme::lme(meanPSD ~ brand + bin + brand * bin, random = ~ 1|serialnumber, 
                                      method="ML", data = df_long$high)
domFreq_intercept_id_interaction_high <- nlme::lme(domFreq ~ brand + bin + brand * bin, random = ~ 1|serialnumber, 
                                      method="ML", data = df_long$high)
```

#### Model assumption: normal distributions of residuals
```{r Mean PSD Model assumptions high sampling frequency experiment, echo=FALSE}
ggpubr::ggqqplot(df_long$high, "meanPSD", facet.by = "bin", 
                 title = "QQ-plot mean power spectral density")
residuals_meanPSD <- resid(meanPSD_intercept_id_high)
#summary(residuals_meanPSD)
hist(residuals_meanPSD)
```

```{r Dominant frequency Model assumptions high sampling frequency experiment, echo=FALSE}
ggpubr::ggqqplot(df_long$high, "domFreq", facet.by = "bin", title = "QQ-plot dominant frequency")
residuals_domFreq <- resid(domFreq_intercept_id_high)
#summary(residuals_domFreq)
hist(residuals_domFreq) # approximately normally distributed
```

#### Results Mean Power Spectral Density

##### Model: meanPSD ~ brand
###### Reference: ActiGraph
```{r Report results for Mean PSD high sampling frequency experiment: ActiGraph, echo=FALSE}
getOutputMM(meanPSD_intercept_id_high)
```
###### Reference: Axivity
```{r Report results for Mean PSD high sampling frequency experiment: Axivity, echo=FALSE}
df_long$high$brand <- relevel(df_long$high$brand, ref = 2) 
meanPSD_intercept_id_high <- nlme::lme(meanPSD ~ brand, random = ~ 1|serialnumber, 
                                       method="ML", data = df_long$high)
getOutputMM(meanPSD_intercept_id_high)
```
##### Model: meanPSD ~ brand + bin + brand*bin
###### Reference: Axivity, bin1
```{r Report results for Mean PSD interaction high sampling frequency experiment: Axivity, echo=FALSE}
getOutputMM(meanPSD_intercept_id_interaction_high)
```

#### Results Dominant Frequency

##### Model: domFreq ~ brand
###### Reference: ActiGraph

```{r Report results for Dominant frequency high sampling frequency experiment: ActiGraph, echo=FALSE}
df_long$high$brand <- relevel(df_long$high$brand, ref = 2) 
domFreq_intercept_id_high <- nlme::lme(domFreq ~ brand, random = ~ 1|serialnumber, 
                                       method="ML", data = df_long$high)
getOutputMM(domFreq_intercept_id_high)
```

##### Reference: Axivity

```{r Report results for Dominant frequency high sampling frequency experiment: Axivity, echo=FALSE}
df_long$high$brand <- relevel(df_long$high$brand, ref = 2) 
domFreq_intercept_id_high <- nlme::lme(domFreq ~ brand, random = ~ 1|serialnumber, 
                                       method="ML", data = df_long$high)
getOutputMM(domFreq_intercept_id_high)
```
##### Model: domFreq ~ brand + bin + brand*bin
###### Reference: Axivity, bin1
```{r Report results for dominant frequency interaction high sampling frequency experiment: Axivity, echo=FALSE}
getOutputMM(domFreq_intercept_id_interaction_high)
```

```{r Create boxplot mean power spectral density, include=FALSE}
gridExtra::grid.arrange(bxp_psd_low, bxp_psd_high, nrow=1) #arranges plots within grid
g <- gridExtra::arrangeGrob(bxp_psd_low + ggplot2::theme(legend.position="none"), 
                            bxp_psd_high + ggplot2::theme(legend.position="bottom"), nrow=2, top = "Low sampling frequency experiment", bottom = "High sampling frequency experiment", left = "Mean power spectral density (g2/Hertz)") #generates g
ggplot2::ggsave(file=paste0(datadir, "plots/boxplot_meanPSD.pdf"), g) #saves g

gridExtra::grid.arrange(bxp_psd_low_perc, bxp_psd_high_perc, nrow=1) #arranges plots within grid
g_perc <- gridExtra::arrangeGrob(bxp_psd_low_perc + ggplot2::theme(legend.position="none"),
                                bxp_psd_high_perc + ggplot2::theme(legend.position="bottom"), 
                                nrow=2, top = "Low sampling frequency experiment", bottom = "High sampling frequency experiment", left = "Mean power spectral density (g2/Hertz)") #generates g_perc
ggplot2::ggsave(file=paste0(datadir, "plots/boxplot_meanPSD_percentage.pdf"), g_perc) #saves g_perc
```


```{r Create boxplot dominant frequency, include=FALSE}
gridExtra::grid.arrange(bxp_domfreq_low, bxp_domfreq_high, nrow=1) #arranges plots within grid
g <- gridExtra::arrangeGrob(bxp_domfreq_low + ggplot2::theme(legend.position="none"), 
                            bxp_domfreq_high + ggplot2::theme(legend.position="bottom"), nrow=2, top = "Low sampling frequency experiment", bottom = "High sampling frequency experiment", left = "Dominant frequency (Hertz)") #generates g
ggplot2::ggsave(file=paste0(datadir, "plots/boxplot_domFreq.pdf"), g) #saves g


gridExtra::grid.arrange(bxp_domfreq_low_perc, bxp_domfreq_high_perc, 
                        nrow=1) #arranges plots within grid
g_perc <- gridExtra::arrangeGrob(bxp_domfreq_low_perc + ggplot2::theme(legend.position="none"),
                                bxp_domfreq_high_perc + ggplot2::theme(legend.position="bottom"), 
                                nrow=2, top = "Low sampling frequency experiment", bottom = "High sampling frequency experiment", left = "Dominant frequency (Hertz)") #generates g_perc
ggplot2::ggsave(file=paste0(datadir, "plots/boxplot_domFreq_percentage.pdf"), g_perc) #saves g_perc
```
