---
title: "Frequency domain analyses"
author: "Annelinde Lettink"
output: 
  html_notebook:
    toc: true
    toc_depth: 4  # upto four depths of headings (specified by #, ##, ###, and ####)
---

```{r Setup and load data, include=FALSE}
#install.packages("knitr")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'pdf')

rm(list=ls())
graphics.off()

# Update to be your local directory, if all goes well this is the only line you will have to update
shaker_experiments_folder = "/Users/annelindelettink/Documents/Work MacBook Pro Annelinde/Mechanical Shaker Machine"
# shaker_experiments_folder = "~/data/VUMC/shaker_experiments"

#====================================================================================
# Specify file paths
datadir = paste0(shaker_experiments_folder, "/analyses/")
if (!dir.exists(datadir)) {
  stop(paste0("Directory does not exist: ", datadir))
}

# Load cleaned mechanical shaker machine data for both experiments
filename_high = paste0(datadir, "ms_flat_HA_cleaned_high.RData") 
filename_low = paste0(datadir, "ms_flat_HA_cleaned_low.RData") 
if (!file.exists(filename_high)) {
  stop(paste0("File does not exist: ", filename_high))
}
load(filename_high)
if (!file.exists(filename_low)) {
  stop(paste0("File does not exist: ", filename_low))
}
load(filename_low)
rm(filename_high, filename_low)
```

```{r Functions, include=FALSE}
#' deriveSpectrum
#'
#' @description 'deriveSpectrum' derives the frequency spectrum of a signal
#'
#' @param x Vector that contains the values of the signal (i.e., acceleration)
#' @param sampling_frequency Integer that indicates the sampling frequency of the signal (i.e., the average number of samples obtained in one second in Hz)
#' @param file_name String to indicate the name of the plot (i.e., "spectrum" result in the following name for plot 1 "spectrum_1.jpeg").
#' @param plot Boolean, default is TRUE the derived spectrum will be plotted
#' @param XLIM Vector that limits the frequency content of the plot, default is c(0, 5) to limit the frequency content to 5 Hz (as 250 rpm / 60 = 4.1667 Hz is the expected max)
#' @return An object of class "spec", which is a list containing at least the following components: 
#' \item{freq}{Vector of frequencies at which the spectral density is estimated.} 
#' \item{spec}{Vector (for univariate series) or matrix (for multivariate series) of estimates of the spectral density at frequencies corresponding to freq.}
#' \item{coh}{NULL for univariate series. For multivariate time series, a matrix containing the squared coherency between different series.}  
#' \item{phase}{NULL for univariate series. For multivariate time series a matrix containing the cross-spectrum phase between different series.} 
#' \item{series}{The name of the time series.}
#' \item{snames}{For multivariate input, the names of the component series.}
#' \item{method}{The method used to calculate the spectrum.} 
#' @export
deriveSpectrum <- function(x, sampling_frequency, file_name, plot = TRUE, XLIM = c(0, 5)) {
  specd <- spectrum(x, log = "no", plot = FALSE) # Default calculate spectrum on log-scale, but we use raw data
  
  # Frequency axis of the spectrum, specd, is calculated in terms of cycles per sampling interval
  delta <- 1/sampling_frequency
  specd$specx <- specd$freq/delta # Converts to cycles per unit time (divide by sampling interval)
  specd$specy <- 2*specd$spec # Multiply the spectrum by 2 to make it actually equal to variance
  
  #Save the frequency spectrum plot and zoom in on low-frequencies (at most 5 based on max rpm)
  if (plot == TRUE){
    jpeg(paste(datadir, paste0(paste(file_name, file, sep = "_"), ".jpeg"), sep = "/plots/"),
         width=600, height=500, res=120) # start export
    plot(specd$specx, specd$specy,
         main = "HA", xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l") 
    dev.off()
  }
  return(specd)
}

#' deriveMeanSpectrum
#'
#' @description 'deriveMeanSpectrum' derives the mean frequency spectrum from the derived spectra
#'
#' @param spectra A list containing the raw spectra over which the mean will be derived
#' @return A list containing two following elements: 
#' \item{x}{Vector of frequencies at which the spectral density (y) is estimated} 
#' \item{y}{Vector of estimates of the spectral density at frequencies corresponding to frequencies (x)}
#' @export
deriveMeanSpectrum <- function(spectra) {
  df_specx <- data.frame()
  df_specy <- data.frame()
  for(spectrum in 1:length(spectra)){
    df_specx <- rbind(df_specx, spectra[[spectrum]]$specx)
    df_specy <- rbind(df_specy, spectra[[spectrum]]$specy)
  }
  x <- colMeans(df_specx)
  y <- colMeans(df_specy)
  spec <- list(x = x, y = y)
  return(spec)
}

#' deriveComparisonValues
#'
#' @description 'deriveComparisonValues' derives the dominant frequency and the mean power spectral density from a frequency spectrum
#'
#' @param spectrum An object of class "spec"
#' @param freqBins Vector containing frequency values that define the frequency bins (can be derived using the function 'defineFreqBins') 
#' @return Data.frame consisting of two columns: \item{domFreq}{The dominant frequency of the signal in each frequency bin} \item{meanPSD}{The mean power spectral density of the signal in each frequency bin}
#' @export
deriveComparisonValues <- function(spectrum, freqBins){
  df <- data.frame()
  for(i in 1:(length(freqBins)-1)){
    lower <- freqBins[i]
    upper <- freqBins[i+1]
    index <- which(spectrum$specx >= lower & spectrum$specx <= upper)
    df[i, 1] <- spectrum$specx[index[which.max(spectrum$specy[index])]] # dominant frequency
    df[i, 2] <- mean(spectrum$specy[index]) # mean power spectral density
  }
  colnames(df) <- c("domFreq", "meanPSD")
  return(df)
}

#' createBoxplot
#'
#' @description 'createBoxplot' creates a boxplot between or within devices for the dominant frequency or the mean power spectral density per frequency bin
#'
#' @param data A data set containing the comparison values for the experiment
#' @param outcome String containing the type of the comparison value. One of c("domFreq", "meanPSD") 
#' @param type String containing the axis the outcome value is derived from. One of c("within", "between") 
#' @param logscale Boolean to indicate whether the vertical axis needs to be logtransformed. One of c("TRUE", "FALSE") 
#' @return Boxplot between or within devices for the outcome (y-axis) per frequency bin (x-axis)
#' @export
createBoxplot <- function(data, outcome = c("domFreq", "meanPSD"), type = c("within", "between"), logscale = c(TRUE, FALSE)){
  if(type == "between"){
    per = "brand"
  } else {
    per = "id"
  }
  if (outcome == "domFreq"){
    label = "Dominant frequency (Hertz)"
  } else{
    label = "Mean power spectral density (g2/Hertz)"
  }
  boxplot <- ggpubr::ggboxplot(data, x = "freqBin", y = outcome, color = per,
                               xlab = "Mean frequency bin (Hertz)", ylab = label, 
                               order = c("t0", "t1", "t2", "t3", "t4", "t5", "t6", 
                                         "t7", "t8", "t9", "t10"), 
                               font.label = list(size = 8, color = "black")) +
    ggpubr::font("xlab", size = 8, color = "black") +
    ggpubr::font("x.text", size = 8, color = "black") + 
    ggpubr::font("ylab", size = 8, color = "black") +
    ggpubr::font("y.text", size = 8, color = "black") + 
    ggpubr::font("legend.title", size = 8, color = "black") +
    ggpubr::font("legend.text", size = 8, color = "black") + 
    ggplot2::scale_x_discrete(breaks=c("t0", "t1", "t2", "t3", "t4", "t5", "t6", "t7", "t8", "t9", "t10"),
                     labels= c("0.25",  "0.67", "1.04", "1.46", "1.88", "2.29", "2.71", "3.13",
                               "3.54", "3.96", "4.59")) 
  if(logscale == TRUE){
    boxplot <- boxplot + ggpubr::yscale("log10", .format = TRUE)
  }
  return(boxplot)
}

#' evaluateMM
#'
#' @description 'evaluateMM' function to indicate if the likelihood ratio test was significant
#'
#' @param model1 Model object with the least parameters
#' @param model2 Model object with the most parameters
#' @param type String containing the addition to the model. One of c("randomIntercept", "randomSlope") 
#' @return Boolean: \item{TRUE}{The addition of a random intercept/slope is necessary} \item{FALSE}{The addition of a random intercept/slope is not necessary}
#' @export
evaluateMM <- function(model1, model2, type = c("randomIntercept", "randomSlope")){
  logLikDiff <- (-2*(logLik(model1))) - (-2*logLik(model2))
  if(type == "randomIntercept"){
    return(logLikDiff > 3.84)
  }
  else{ 
    return(logLikDiff > 5.99)
  }
}

#' getOutputMM
#'
#' @description 'getOutputMM' function print the model output
#'
#' @param finalModel Model object for which the output will be printed
#' @return Printed model output including fixed and random effect estimates
#' @export
getOutputMM <- function(finalModel) {
  print(summary(finalModel)) #fixed effect estimates
  lme4::VarCorr(finalModel) #random effect estimates
}
```

# Frequency spectra

This section presents the frequency spectra. First the mean frequency spectra for both experiments plotted to derive frequency bins. Second, frequency spectra are printed for each accelerometer recording including red lines that indicate these bins for both experiments seperately.

```{r Settings for frequency spectra, include=FALSE}
# Plot parameters
plot = TRUE
xlabel = "Frequency (Hertz)"
ylabel = "Spectral Density (g2/Hertz)"
XLIM = c(0, 5) # Limit frequency content to 5 Hz (as 250 rpm / 60 = 4.1667 Hz is the expected max)
```

## Mean frequency spectra
### Low sampling frequency experiment

```{r Derive frequency spectra low sampling frequency experiment, echo=FALSE}
# DERIVE RAW FREQUENCY SPECTRA (for visual inspection and save data for further analysis)
raw.spectra_low <- list() 

for (file in 1:length(clean_data_low$data)) { # for all accelerometer files
  sampling_frequency <- as.numeric(clean_data_low$specifications$sampling_frequency[file])  
  HA <- clean_data_low$data[[file]]$HA #for horizontal axis
  rawSpectrum <- deriveSpectrum(HA, sampling_frequency, file_name = names(clean_data_low$data)[[file]])
  raw.spectra_low[[file]] <- rawSpectrum
}
rm(rawSpectrum, file)
names(raw.spectra_low) <- clean_data_low$specifications$label  

mean.freqspec_low <- deriveMeanSpectrum(raw.spectra_low)
```

```{r Plot mean frequency spectrum low sampling frequency, echo=FALSE}
plot(mean.freqspec_low$x, mean.freqspec_low$y, xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l", main = "Mean frequency spectrum low sampling frequency experiment")
```

### High sampling frequency experiment

```{r Derive frequency spectra high sampling frequency experiment, echo=FALSE}
# DERIVE RAW FREQUENCY SPECTRA (for visual inspection and save data for further analysis)
raw.spectra_high <- list() 

for (file in 1:length(clean_data_high$data)) { # for all accelerometer files
  sampling_frequency <- as.numeric(clean_data_high$specifications$sampling_frequency[file])  
  HA <- clean_data_high$data[[file]]$HA #for horizontal axis
  rawSpectrum <- deriveSpectrum(HA, sampling_frequency, file_name = names(clean_data_high$data)[[file]])
  raw.spectra_high[[file]] <- rawSpectrum
}
rm(rawSpectrum, file)
names(raw.spectra_high) <- clean_data_high$specifications$label

mean.freqspec_high <- deriveMeanSpectrum(raw.spectra_high)
```

```{r Plot mean frequency spectrum high sampling frequency, echo=FALSE}
plot(mean.freqspec_high$x, mean.freqspec_high$y, xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l", main = "Mean frequency spectrum high sampling frequency experiment")
```

```{r Save raw spectra, include=FALSE}
raw_spectra <- list(high = raw.spectra_high, high_mean = mean.freqspec_high, specifications_high = clean_data_high$specifications,
                    low = raw.spectra_low, low_mean = mean.freqspec_low, specifications_low = clean_data_low$specifications) # Save derived raw spectra for HA in one list
#cat("\nSaving data...")
save(raw_spectra, file = paste0(datadir, "frequency_spectra.RData"))
rm(raw.spectra_high, raw.spectra_low, HA, sampling_frequency)
```

## Frequency bins
We grouped the 19 shaker conditions (30, 40, 50, 62, 75, 87, 100, 112, 125, 137, 150, 160, 175, 187, 200, 212, 225, 237, and 250 rotations per minute) into groups of two, and divided the boundary values between the transitions (56, 81, 106, 131, 155, 181, 206, 231, and <300 rpm) by 60 to derive the associated frequency in Hertz. This resulted in the following associated frequencies: 0.93, 1.35, 1.77, 2.18, 2.58, 3.02, 3.43, 3.85, and 5.00 Hz. 
These associated frequencies were used as cut-offs for frequency bins, over which aggregated outcome values (i.e., dominant frequency and mean power spectral density) were calculated for statistical comparison between brands.

```{r Derive frequency bins, include=FALSE}
rpm <- c(56, 81, 106, 131, 155, 181, 206, 231)
shaking_frequencies <- rpm / 60
rm(rpm)
```

### Frequency spectrum plots
This section presents the frequency spectrum plots of the device acceleration signals included in the analyses with the frequency bins indicated by red vertical lines.

#### High sampling frequency experiment

```{r Plot spectra high sampling frequency experiment, echo=FALSE}
#load(paste0(datadir, "frequency_spectra.RData"))

for (spec in 1:length(raw_spectra$high)) {
  spectrum <- raw_spectra$high[[spec]]
  title <- paste("Frequency spectrum plot: ", raw_spectra$specifications_high$label[[spec]])
  plot(spectrum$specx, spectrum$specy, main = title, 
       xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l") 
  for(i in 1:length(shaking_frequencies)){
    abline(v=shaking_frequencies[i], col = "red")
    abline(v=shaking_frequencies[i], col = "red")
  }  
}
rm(spec, title, spectrum, i)
```

#### Low sampling frequency experiment

```{r Plot spectra low sampling frequency experiment, echo=FALSE}
for (spec in 1:length(raw_spectra$low)) {
  spectrum <- raw_spectra$low[[spec]]
  title <- paste("Frequency spectrum plot: ", raw_spectra$specifications_low$label[[spec]])
  plot(spectrum$specx, spectrum$specy, main = title, 
       xlab = xlabel, ylab = ylabel, type = "l", xlim = XLIM, bty= "l")
  for(i in 1:length(shaking_frequencies)){
    abline(v=shaking_frequencies[i], col = "red")
    abline(v=shaking_frequencies[i], col = "red")
  }  
}
rm(spec, title, spectrum, i)
```

```{r Compute dominant frequency (domFreq) and mean power spectral density (meanPSD) for HA , include = FALSE}
# Derive frequency bins from the cutoffs
freqBins <- round(c(XLIM[1], shaking_frequencies, XLIM[2]), digits = 2)
rm(rpm, shaking_frequencies)

comparison_values_low <- list()
for(spectrum in 1:length(raw_spectra$low)){ 
  comparison_values_low[[spectrum]] <- deriveComparisonValues(raw_spectra$low[[spectrum]], freqBins)
}
names(comparison_values_low) <- names(raw_spectra$low)

comparison_values_high <- list()
for(spectrum in 1:length(raw_spectra$high)){ 
  comparison_values_high[[spectrum]] <- deriveComparisonValues(raw_spectra$high[[spectrum]], freqBins)
}
names(comparison_values_high) <- names(raw_spectra$high)
#cat("\nSaving data...")
comparison_values <- list(high = comparison_values_high, low = comparison_values_low)
save(comparison_values, file = paste0(datadir, "domFreq_meanPSD.RData")) 

rm(spectrum, comparison_values_high, comparison_values_low)
```

```{r WIDE FORMAT: Prepare data set for statistical analyses, include = FALSE}
# Wide format: for both experiments and outcomes domFreq and meanPSD

df_wide_low <- data.frame()
id <- names(comparison_values$low)
df_wide_low <- cbind(id, clean_data_low$specifications$brand, clean_data_low$specifications$sampling_frequency,
                 clean_data_low$specifications$dynamic_range)
domFreq_wide_low <- data.frame() 
meanPSD_wide_low <- data.frame()
for (accelerometer in 1:length(comparison_values$low)) {
  domFreq_wide_low <- rbind(domFreq_wide_low, comparison_values$low[[accelerometer]]$domFreq)
  meanPSD_wide_low <- rbind(meanPSD_wide_low, comparison_values$low[[accelerometer]]$meanPSD)
}
domFreq_ms_wide_low <- cbind(df_wide_low, domFreq_wide_low)
colnames(domFreq_ms_wide_low) <- c("id", "brand", "sampling_frequency", "dynamic_range", 
                               "t0", "t1", "t2", "t3", "t4", "t5", "t6", "t7", "t8")
meanPSD_ms_wide_low <- cbind(df_wide_low, meanPSD_wide_low)
colnames(meanPSD_ms_wide_low) <- c("id", "brand", "sampling_frequency", "dynamic_range", 
                               "t0", "t1", "t2", "t3", "t4", "t5", "t6", "t7", "t8")
rm(df_wide_low, id, domFreq_wide_low, meanPSD_wide_low, accelerometer)

df_wide_high <- data.frame()
id <- names(comparison_values$high)
df_wide_high <- cbind(id, clean_data_high$specifications$brand, clean_data_high$specifications$sampling_frequency,
                 clean_data_high$specifications$dynamic_range)
domFreq_wide_high <- data.frame() 
meanPSD_wide_high <- data.frame()
for (accelerometer in 1:length(comparison_values$high)) {
  domFreq_wide_high <- rbind(domFreq_wide_high, comparison_values$high[[accelerometer]]$domFreq)
  meanPSD_wide_high <- rbind(meanPSD_wide_high, comparison_values$high[[accelerometer]]$meanPSD)
}
domFreq_ms_wide_high <- cbind(df_wide_high, domFreq_wide_high)
colnames(domFreq_ms_wide_high) <- c("id", "brand", "sampling_frequency", "dynamic_range", 
                               "t0", "t1", "t2", "t3", "t4", "t5", "t6", "t7", "t8")
meanPSD_ms_wide_high <- cbind(df_wide_high, meanPSD_wide_high)
colnames(meanPSD_ms_wide_high) <- c("id", "brand", "sampling_frequency", "dynamic_range", 
                               "t0", "t1", "t2", "t3", "t4", "t5", "t6", "t7", "t8")
rm(df_wide_high, id, domFreq_wide_high, meanPSD_wide_high, accelerometer)
```

```{r LONG FORMAT: Prepare data set for statistical analyses, include = FALSE}
# Combining domFreq and meanPSD (gather columns for outcome variables)
library(tidyverse)

domFreq_long_low <- domFreq_ms_wide_low %>%
  gather(key = "freqBin", value = "domFreq", t0, t1, t2, t3, t4, t5, t6, t7, t8) %>%
  rstatix::convert_as_factor(id, freqBin, brand, sampling_frequency, dynamic_range)
meanPSD_long_low <- meanPSD_ms_wide_low %>%
  gather(key = "freqBin", value = "meanPSD", t0, t1, t2, t3, t4, t5, t6, t7, t8) %>%
  rstatix::convert_as_factor(id, freqBin, brand, sampling_frequency, dynamic_range)
df_low <- merge(domFreq_long_low, meanPSD_long_low) # Join both data.frame objects to derive a single dataset
rm(domFreq_long_low, domFreq_ms_wide_low, meanPSD_long_low, meanPSD_ms_wide_low)
df_low <- df_low[order(df_low$id),]

domFreq_long_high <- domFreq_ms_wide_high %>%
  gather(key = "freqBin", value = "domFreq", t0, t1, t2, t3, t4, t5, t6, t7, t8) %>%
  rstatix::convert_as_factor(id, freqBin, brand, sampling_frequency, dynamic_range)
meanPSD_long_high <- meanPSD_ms_wide_high %>%
  gather(key = "freqBin", value = "meanPSD", t0, t1, t2, t3, t4, t5, t6, t7, t8) %>%
  rstatix::convert_as_factor(id, freqBin, brand, sampling_frequency, dynamic_range)
df_high <- merge(domFreq_long_high, meanPSD_long_high) # Join both data.frame objects to derive a single dataset
rm(domFreq_long_high, domFreq_ms_wide_high, meanPSD_long_high, meanPSD_ms_wide_high)
df_high <- df_high[order(df_high$id),]

df_long <- list(low = df_low, high = df_high)
#cat("\nSaving data...")
save(df_long, file = paste0(datadir, "domfreq_meanPSD_long.RData"))
rm(df_high, df_low)
```

# Statistical comparison

This section presents the statistical comparison of the frequency spectra in terms of the aggregated dominant frequency and mean power spectral density for each frequency bin per sampling frequency experiment.

## Low sampling frequency experiment

```{r Create boxplots low sampling frequency experiment, include=FALSE}
#load(paste0(datadir, "domfreq_meanPSD_long.RData"))

bxp_domfreq_low <- createBoxplot(df_long$low, outcome = "domFreq", type = "between", logscale = FALSE)
bxp_psd_low <- createBoxplot(df_long$low, outcome = "meanPSD", type = "between", logscale = FALSE)
gridExtra::grid.arrange(bxp_domfreq_low, bxp_psd_low, nrow=1) #arranges plots within grid
g <- gridExtra::arrangeGrob(bxp_domfreq_low + ggplot2::theme(legend.position="none"), 
                            bxp_psd_low + ggplot2::theme(legend.position="bottom"), nrow=2) #generates g 
ggplot2::ggsave(file=paste0(datadir, "plots/boxplots_HA_LOW.pdf"), g) #saves g

bxp_domfreq_low_log <- createBoxplot(df_long$low, outcome = "domFreq", type = "between", 
                                     logscale = TRUE)
bxp_psd_low_log <- createBoxplot(df_long$low, outcome = "meanPSD", type = "between", logscale = TRUE) 
gridExtra::grid.arrange(bxp_domfreq_low_log, bxp_psd_low_log, nrow=1) #arranges plots within grid
g_log <- gridExtra::arrangeGrob(bxp_domfreq_low_log + ggplot2::theme(legend.position="none"),
                                bxp_psd_low_log + ggplot2::theme(legend.position="bottom"),
                                nrow=2)#generates g_log
ggplot2::ggsave(file=paste0(datadir, "plots/boxplots_HA_LOW_logscale.pdf"), g_log) #saves g_log
```

### Boxplots
This section presents the boxplots for both outcomes (dominant frequency and mean power spectral density) for the low sampling frequency experiment.

Boxplot
```{r Print boxplot low sampling frequency experiment, echo=FALSE}
plot(g)
```

Boxplot with log scale
```{r Print log boxplot low sampling frequency experiment, echo=FALSE}
plot(g_log)
```

### Mixed model analyses
This section presents the statistical comparison between the brand for the low sampling frequency experiment. Thee mixed models for mean power spectral density and dominant frequency with a random intercept for serial number (id) are presented separately.

```{r Mixed Model Analyses low sampling frequency experiment, include=FALSE}
# Starting model: with random intercept for accelerometer id (to adjust for the repeated measurements of the accelerometer devices)
meanPSD_intercept_id_low <- nlme::lme(meanPSD ~ brand, random = ~ 1|id, 
                                      method="ML", data = df_long$low)
domFreq_intercept_id_low <- nlme::lme(domFreq ~ brand, random = ~ 1|id, 
                                      method="ML", data = df_long$low)

# Add random intercept for freqBin (to adjust for the frequencies)
meanPSD_intercept_freqbin_low <- nlme::lme(meanPSD ~ brand, random = list(~ 1|id, ~ 1|freqBin),
                                           method="ML", data = df_long$low)
domFreq_intercept_freqbin_low <- nlme::lme(domFreq ~ brand, random = list(~ 1|id, ~ 1|freqBin),
                                           method="ML", data = df_long$low)

# Evaluate if adding a random intercept for freqBin is necessary
evaluateMM(meanPSD_intercept_id_low, meanPSD_intercept_freqbin_low, type = "randomIntercept") #FALSE adding a random intercept for frequency bin to the model is not necessary
evaluateMM(domFreq_intercept_id_low, domFreq_intercept_freqbin_low, type = "randomIntercept") #FALSE adding a random intercept for frequency bin to the model is not necessary
rm(meanPSD_intercept_freqbin_low, domFreq_intercept_freqbin_low)
```

#### Model assumption: normal distribution of residuals

```{r Mean PSD Model assumptions low sampling frequency experiment, echo=FALSE}
ggpubr::ggqqplot(df_long$low, "meanPSD", facet.by = "freqBin", title = "QQ-plot mean power spectral density")
residuals_meanPSD <- resid(meanPSD_intercept_id_low)
#summary(residuals_meanPSD)
hist(residuals_meanPSD)
```

```{r Dominant frequency Model assumptions low sampling frequency experiment, echo=FALSE}
ggpubr::ggqqplot(df_long$low, "domFreq", facet.by = "freqBin", title = "QQ-plot dominant frequency")
residuals_domFreq <- resid(domFreq_intercept_id_low)
#summary(residuals_domFreq)
hist(residuals_domFreq) # approximately normally distributed
```

#### Results for the Mean Power Spectral Density

##### Reference: ActiGraph 
```{r Report results for Mean PSD low sampling frequency experiment: ActiGraph, echo=FALSE}
getOutputMM(meanPSD_intercept_id_low)
```
##### Reference: MOX
```{r Report results for Mean PSD low sampling frequency experiment: MOX, echo=FALSE}
df_long$low$brand <- relevel(df_long$low$brand, ref = 5) 
meanPSD_intercept_id_low <- nlme::lme(meanPSD ~ brand, random = ~ 1|id, 
                                      method="ML", data = df_long$low)
getOutputMM(meanPSD_intercept_id_low)
```
##### Reference: GENEActiv
```{r Report results for Mean PSD low sampling frequency experiment: GENEActiv, echo=FALSE}
df_long$low$brand <- relevel(df_long$low$brand, ref = 5) 
meanPSD_intercept_id_low <- nlme::lme(meanPSD ~ brand, random = ~ 1|id, 
                                      method="ML", data = df_long$low)
getOutputMM(meanPSD_intercept_id_low)
```
##### Reference: Axivity
```{r Report results for Mean PSD low sampling frequency experiment: Axivity, echo=FALSE}
df_long$low$brand <- relevel(df_long$low$brand, ref = 5) 
meanPSD_intercept_id_low <- nlme::lme(meanPSD ~ brand, random = ~ 1|id, 
                                      method="ML", data = df_long$low)
getOutputMM(meanPSD_intercept_id_low)
```

##### Reference: activPAL
```{r Report results for Mean PSD low sampling frequency experiment: activPAL, echo=FALSE}
df_long$low$brand <- relevel(df_long$low$brand, ref = 5) 
meanPSD_intercept_id_low <- nlme::lme(meanPSD ~ brand, random = ~ 1|id, 
                                      method="ML", data = df_long$low)
getOutputMM(meanPSD_intercept_id_low)
```
#### Results for the Dominant Frequency

##### Reference: ActiGraph
```{r Report results for Dominant frequency low sampling frequency experiment: ActiGraph, echo=FALSE}
df_long$low$brand <- relevel(df_long$low$brand, ref = 5) 
domFreq_intercept_id_low <- nlme::lme(domFreq ~ brand, random = ~ 1|id, 
                                      method="ML", data = df_long$low)
getOutputMM(domFreq_intercept_id_low)
```
##### Reference: MOX
```{r Report results for Dominant frequency low sampling frequency experiment: MOX, echo=FALSE}
df_long$low$brand <- relevel(df_long$low$brand, ref = 5) 
domFreq_intercept_id_low <- nlme::lme(domFreq ~ brand, random = ~ 1|id, 
                                      method="ML", data = df_long$low)
getOutputMM(domFreq_intercept_id_low)
```
##### Reference: GENEActiv
```{r Report results for Dominant frequency low sampling frequency experiment: GENEActiv, echo=FALSE}
df_long$low$brand <- relevel(df_long$low$brand, ref = 5) 
domFreq_intercept_id_low <- nlme::lme(domFreq ~ brand, random = ~ 1|id, 
                                      method="ML", data = df_long$low)
getOutputMM(domFreq_intercept_id_low)
```
##### Reference: Axivity

```{r Report results for Dominant frequency low sampling frequency experiment: Axivity, echo=FALSE}
df_long$low$brand <- relevel(df_long$low$brand, ref = 5) 
domFreq_intercept_id_low <- nlme::lme(domFreq ~ brand, random = ~ 1|id, 
                                      method="ML", data = df_long$low)
getOutputMM(domFreq_intercept_id_low)
```
##### Reference: activPAL

```{r Report results for Dominant frequency low sampling frequency experiment: activPAL, echo=FALSE}
df_long$low$brand <- relevel(df_long$low$brand, ref = 5) 
domFreq_intercept_id_low <- nlme::lme(domFreq ~ brand, random = ~ 1|id, 
                                      method="ML", data = df_long$low)
getOutputMM(domFreq_intercept_id_low)
```

## High sampling frequency experiment

### Boxplots
This section presents the boxplots for both outcomes (dominant frequency and mean power spectral density) for the high sampling frequency experiment.

```{r Create boxplots high sampling frequency, include=FALSE}
#load(paste0(datadir, "domfreq_meanPSD_ms_HA_dataset_long_hf.RData"))
bxp_domfreq_high <- createBoxplot(df_long$high, outcome = "domFreq", type = "between", 
                                  logscale = FALSE)
bxp_psd_high <- createBoxplot(df_long$high, outcome = "meanPSD", type = "between", logscale = FALSE) 

gridExtra::grid.arrange(bxp_domfreq_high, bxp_psd_high, nrow=1) #arranges plots within grid
g <- gridExtra::arrangeGrob(bxp_domfreq_high + ggplot2::theme(legend.position="none"), 
                            bxp_psd_high + ggplot2::theme(legend.position="bottom"), nrow=2) #generates g
ggplot2::ggsave(file=paste0(datadir, "plots/boxplots_HA_HIGH.pdf"), g) #saves g

bxp_domfreq_high_log <- createBoxplot(df_long$high, outcome = "domFreq", type = "between", 
                                      logscale = TRUE)
bxp_psd_high_log <- createBoxplot(df_long$high, outcome = "meanPSD", type = "between", logscale = TRUE)
gridExtra::grid.arrange(bxp_domfreq_high_log, bxp_psd_high_log, 
                        nrow=1) #arranges plots within grid
g_log <- gridExtra::arrangeGrob(bxp_domfreq_high_log + ggplot2::theme(legend.position="none"),
                                bxp_psd_high_log + ggplot2::theme(legend.position="bottom"), 
                                nrow=2) #generates g_log
ggplot2::ggsave(file=paste0(datadir, "plots/boxplots_HA_HIGH_logscale.pdf"), g_log) #saves g_log
```

Boxplot

```{r Print boxplot high sampling frequency experiment, echo=FALSE}
plot(g)
```

Boxplot with log scale

```{r Print log boxplot high sampling frequency experiment, echo=FALSE}
plot(g_log)
```

### Mixed model analyses
This section presents the statistical comparison between the brand for the high sampling frequency experiment. Thee mixed models for mean power spectral density and dominant frequency with a random intercept for serial number (id) are presented separately.

```{r Mixed Model Analyses high sampling frequency experiment, include=FALSE}
# Starting model: with random intercept for accelerometer id (to adjust for the repeated measurements of the accelerometer devices)
meanPSD_intercept_id_high <- nlme::lme(meanPSD ~ brand, random = ~ 1|id, 
                                       method="ML", data = df_long$high)
domFreq_intercept_id_high <- nlme::lme(domFreq ~ brand, random = ~ 1|id, 
                                       method="ML", data = df_long$high)

# Add random intercept for freqBin (to adjust for the frequencies)
meanPSD_intercept_freqbin_high <- nlme::lme(meanPSD ~ brand, random = list(~ 1|id, ~ 1|freqBin), 
                                            method="ML", data = df_long$high)
domFreq_intercept_freqbin_high <- nlme::lme(domFreq ~ brand, random = list(~ 1|id, ~ 1|freqBin), 
                                            method="ML", data = df_long$high)

# Evaluate if adding a random intercept for freqBin is necessary
evaluateMM(meanPSD_intercept_id_high, meanPSD_intercept_freqbin_high, type = "randomIntercept") #FALSE adding a random intercept for frequency bin to the model is not necessary
evaluateMM(domFreq_intercept_id_high, domFreq_intercept_freqbin_high, type = "randomIntercept") #FALSE adding a random intercept for frequency bin to the model is not necessary
rm(meanPSD_intercept_freqbin_high, domFreq_intercept_freqbin_high)
```

#### Model assumption: normal distributions of residuals
```{r Mean PSD Model assumptions high sampling frequency experiment, echo=FALSE}
ggpubr::ggqqplot(df_long$high, "meanPSD", facet.by = "freqBin", 
                 title = "QQ-plot mean power spectral density")
residuals_meanPSD <- resid(meanPSD_intercept_id_high)
#summary(residuals_meanPSD)
hist(residuals_meanPSD)
```

```{r Dominant frequency Model assumptions high sampling frequency experiment, echo=FALSE}
ggpubr::ggqqplot(df_long$high, "domFreq", facet.by = "freqBin", title = "QQ-plot dominant frequency")
residuals_domFreq <- resid(domFreq_intercept_id_high)
#summary(residuals_domFreq)
hist(residuals_domFreq) # approximately normally distributed
```

#### Results for the Mean Power Spectral Density

##### Reference: ActiGraph
```{r Report results for Mean PSD high sampling frequency experiment: ActiGraph, echo=FALSE}
getOutputMM(meanPSD_intercept_id_high)
```
##### Reference: Axivity
```{r Report results for Mean PSD high sampling frequency experiment: Axivity, echo=FALSE}
df_long$high$brand <- relevel(df_long$high$brand, ref = 2) 
meanPSD_intercept_id_high <- nlme::lme(meanPSD ~ brand, random = ~ 1|id, 
                                       method="ML", data = df_long$high)
getOutputMM(meanPSD_intercept_id_high)
```

#### Results for the Dominant Frequency

##### Reference: Axivity

```{r Report results for Dominant frequency high sampling frequency experiment: Axivity, echo=FALSE}
df_long$high$brand <- relevel(df_long$high$brand, ref = 2) 
domFreq_intercept_id_high <- nlme::lme(domFreq ~ brand, random = ~ 1|id, 
                                       method="ML", data = df_long$high)
getOutputMM(domFreq_intercept_id_high)
```

##### Reference: ActiGraph

```{r Report results for Dominant frequency high sampling frequency experiment: ActiGraph, echo=FALSE}
df_long$high$brand <- relevel(df_long$high$brand, ref = 2) 
domFreq_intercept_id_high <- nlme::lme(domFreq ~ brand, random = ~ 1|id, 
                                       method="ML", data = df_long$high)
getOutputMM(domFreq_intercept_id_high)
```


```{r Create boxplot mean power spectral density, include=FALSE}
gridExtra::grid.arrange(bxp_psd_low, bxp_psd_high, nrow=1) #arranges plots within grid
g <- gridExtra::arrangeGrob(bxp_psd_low + ggplot2::theme(legend.position="none"), 
                            bxp_psd_high + ggplot2::theme(legend.position="bottom"), nrow=2, top = "Low sampling frequency experiment", bottom = "High sampling frequency experiment", left = "Mean power spectral density (g2/Hertz)") #generates g
ggplot2::ggsave(file=paste0(datadir, "plots/boxplot_meanPSD.pdf"), g) #saves g

gridExtra::grid.arrange(bxp_psd_low_log, bxp_psd_high_log, nrow=1) #arranges plots within grid
g_log <- gridExtra::arrangeGrob(bxp_psd_low_log + ggplot2::theme(legend.position="none"),
                                bxp_psd_high_log + ggplot2::theme(legend.position="bottom"), 
                                nrow=2, top = "Low sampling frequency experiment", bottom = "High sampling frequency experiment", left = "Mean power spectral density (g2/Hertz)") #generates g_log
ggplot2::ggsave(file=paste0(datadir, "plots/boxplot_meanPSD_log.pdf"), g_log) #saves g_log
```


```{r Create boxplot dominant frequency, include=FALSE}
gridExtra::grid.arrange(bxp_domfreq_low, bxp_domfreq_high, nrow=1) #arranges plots within grid
g <- gridExtra::arrangeGrob(bxp_domfreq_low + ggplot2::theme(legend.position="none"), 
                            bxp_domfreq_high + ggplot2::theme(legend.position="bottom"), nrow=2, top = "Low sampling frequency experiment", bottom = "High sampling frequency experiment", left = "Dominant frequency (Hertz)") #generates g
ggplot2::ggsave(file=paste0(datadir, "plots/boxplot_domFreq.pdf"), g) #saves g


gridExtra::grid.arrange(bxp_domfreq_low_log, bxp_domfreq_high_log, 
                        nrow=1) #arranges plots within grid
g_log <- gridExtra::arrangeGrob(bxp_domfreq_low_log + ggplot2::theme(legend.position="none"),
                                bxp_domfreq_high_log + ggplot2::theme(legend.position="bottom"), 
                                nrow=2, top = "Low sampling frequency experiment", bottom = "High sampling frequency experiment", left = "Dominant frequency (Hertz)") #generates g_log
ggplot2::ggsave(file=paste0(datadir, "plots/boxplot_domFreq_log.pdf"), g_log) #saves g_log
```
