---
title: "Visual data inspection"
author: "Annelinde Lettink"
output: 
  html_notebook:
    toc: true
    toc_depth: 3 # upto three depths of headings (specified by #, ##, and ###)
---

This file presents the data from the five mechanical shaker experiments: ms_hfcr, ms_lfcr, ms_hfmr, ms_lfmr, ms_bag. Raw acceleration signals of each device per experiment are plotted for visual inspection.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
graphics.off()

#==============================================================================
# Specify file paths and load data
shaker_experiments_folder = "/Users/annelindelettink/Documents/Work MacBook Pro Annelinde/Mechanical Shaker Machine"

datadir = paste0(shaker_experiments_folder, "/analyses/")
if (!dir.exists(datadir)) {
  stop(paste0("Directory does not exist: ", datadir))
}
# Load mechanical shaker data for all five experiments
filename_five_experiments = paste0(datadir, "complete_data.RData") 
if (!file.exists(filename_five_experiments)) {
  stop(paste0("File does not exist: ", filename_five_experiments))
}
load(filename_five_experiments)
```

```{r Functions, include=FALSE}
#' plotSignals
#'
#' @description 'plotSignals' plots the signals in the data set
#'
#' @param outputdir A string that indicates the file path of folder (where the signal plot will be stored); not stored yet
#' @param data List of elements representing a data.frame for one signal, consisting of: 
#' \item{time}{Time stamps}
#' \item{x}{Acceleration signal of the x-axis}
#' \item{y}{Acceleration signal of the y-axis}
#' \item{z}{Acceleration signal of the z-axis}
#' \item{shaking_frequency}{Indicating the movement applied by the shaker machine in rpm}
#' \item{condition}{Indicating the attatchment of the accelerometers}
#' @param axes A vector of characters, one of c("x", "y", "z"), to indicate which axes are required for signal plotting
#' @return A list with the signal plots
#' 
#' @export
#' 
plotSignals <- function(outputdir, d, axes){
  plot_list <- list()
  for(signal in 1:length(d)) {
    file_name <- paste0("acceleration data: ", names(d)[signal])

    jpeg(paste0(outputdir, paste0(file_name, ".jpeg")), width=600, height=500, res=120) 
    colors  <- c( "x-axis" = "red", "y-axis" = "black", "z-axis" = "green")

    p <- ggplot2::ggplot()
    for(axis in 1:length(axes)){
      if(axes[axis] == "x"){
        p <- p + ggplot2::geom_point(data = d[[signal]], ggplot2::aes(x = time, y = x, color = "x-axis") , alpha = 0.2, size = 0.25)
      } else if(axes[axis] == "y"){
        p <- p + ggplot2::geom_point(data = d[[signal]], ggplot2::aes(x = time, y = y, color = "y-axis"), alpha = 0.2, size = 0.25) 
      } else if(axes[axis] == "z"){
        p <- p + ggplot2::geom_point(data = d[[signal]], ggplot2::aes(x = time, y = z, color = "z-axis"), alpha = 0.2, size = 0.25) }
    }
    
    
    p <- p + ggplot2::theme_bw() +
      ggplot2::labs(title = file_name, y = "acceleration (g)", color = "Axis") +
      ggplot2::scale_color_manual(values = colors)
    
    p
    dev.off()
    plot_list[[signal]] <- p
  }
  return(plot_list)
}
```

# Five mechanical shaker experiments
For each device we plotted and inspected the raw acceleration data separately for each experiment.

## Experiment 1 (E1): Low sampling frequency and mixed dynamic range (lfmr)
```{r Signal plots ms_lfmr, include = FALSE}
index <- which(data$specifications$experiment == "ms_lfmr")
ms_lfmr <- list()
for (signal in 1:length(index)){
  ms_lfmr[[signal]] <- data$data[[index[signal]]]
}
names(ms_lfmr) <- data$specifications$label[index]

# Save a plot for each signal
plots.ms_lfmr <- plotSignals(outputdir = paste0(datadir, "plots"), d = ms_lfmr, axes = c("x", "y", "z"))
```

```{r Plot signals ms_lfmr, echo = FALSE}
# Save and print the plot for each signal
for(plot in 1:length(plots.ms_lfmr)){
  print(plots.ms_lfmr[[plot]])
}
rm(plot, plots.ms_lfmr)
```

## Experiment 2 (E2): High sampling frequency and mixed dynamic range (hfmr)
```{r Signal plots ms_hfmr, include = FALSE}
index <- which(data$specifications$experiment == "ms_hfmr")
ms_hfmr <- list()
for (signal in 1:length(index)){
  ms_hfmr[[signal]] <- data$data[[index[signal]]]
}
names(ms_hfmr) <- data$specifications$label[index]

# Save a plot for each signal
plots.ms_hfmr <- plotSignals(outputdir = paste0(datadir, "plots"), d = ms_hfmr, axes = c("x", "y", "z"))
```

```{r Plot signals ms_hfmr, echo = FALSE}
# Save and print the plot for each signal
for(plot in 1:length(plots.ms_hfmr)){
  print(plots.ms_hfmr[[plot]])
}
rm(plot, plots.ms_hfmr)
```

## Experiment 3 (E3): Low sampling frequency and constant dynamic range (lfcr)
Visual inspection shows no typical shaker movements for acceleration signals of aP_490 and aP_254.

```{r Signal plots ms_lfcr, include = FALSE}
index <- which(data$specifications$experiment == "ms_lfcr")
ms_lfcr <- list()
for (signal in 1:length(index)){
  ms_lfcr[[signal]] <- data$data[[index[signal]]]
}
names(ms_lfcr) <- data$specifications$label[index]

# Save a plot for each signal
plots.ms_lfcr <- plotSignals(outputdir = paste0(datadir, "plots"), d = ms_lfcr, axes = c("x", "y", "z"))
```

```{r Plot signals ms_lfcr, echo = FALSE}
# Save and print the plot for each signal
for(plot in 1:length(plots.ms_lfcr)){
  print(plots.ms_lfcr[[plot]])
}
rm(plot, plots.ms_lfcr)
```

## Experiment 4 (E4): High sampling frequency and constant dynamic range (hfcr)
Note that we restarted this experiment, as we noticed that one GENEActiv device (GA_301) had come loose. MOX devices quit recording around 30 minutes before the end of the experiment.

```{r Signal plots ms_hfcr, include = FALSE}
index <- which(data$specifications$experiment == "ms_hfcr")
ms_hfcr <- list()
for (signal in 1:length(index)){
  ms_hfcr[[signal]] <- data$data[[index[signal]]]
}
names(ms_hfcr) <- data$specifications$label[index]

# Save a plot for each signal
plots.ms_hfcr <- plotSignals(outputdir = paste0(datadir, "plots"), d = ms_hfcr, axes = c("x", "y", "z"))
```

```{r Plot signals ms_hfcr, echo = FALSE}
# Save and print the plot for each signal
for(plot in 1:length(plots.ms_hfcr)){
  print(plots.ms_hfcr[[plot]])
}
rm(plot, plots.ms_hfcr)
```

## Experiment 5 (E5): High sampling frequency and lack of standardized orientation (bag)

```{r Signal plots ms_bag, include = FALSE}
index <- which(data$specifications$experiment == "ms_bag")
ms_bag <- list()
for (signal in 1:length(index)){
  ms_bag[[signal]] <- data$data[[index[signal]]]
}
names(ms_bag) <- data$specifications$label[index]

# Save a plot for each signal
plots.ms_bag <- plotSignals(outputdir = paste0(datadir, "plots"), d = ms_bag, axes = c("x", "y", "z"))
```

```{r Plot signals ms_bag, echo = FALSE}
# Save and print the plot for each signal
for(plot in 1:length(plots.ms_bag)){
  print(plots.ms_bag[[plot]])
}
rm(plot, plots.ms_bag)
```
