---
title: "Visual data inspection"
author: "Annelinde Lettink"
output:
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
  html_notebook:
    toc: yes
    toc_depth: 3
---

# Five mechanical shaker experiments

This file presents the data from the five mechanical shaker experiments: E1: ms\_lrcr, E2: ms\_hrcr, E3: ms\_lrmr, E4: ms\_hrmr, E5: ms\_mrcr, and bag: ms\_bag. We plotted raw triaxial acceleration signals of each device per experiment for visual inspection.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
graphics.off()

#==============================================================================
# Specify file paths and load data
shaker_experiments_folder = "/Users/annelindelettink/Documents/Work MacBook Pro Annelinde/Mechanical Shaker Machine"

datadir = paste0(shaker_experiments_folder, "/subsetted_raw_data/")
if (!dir.exists(datadir)) {
  stop(paste0("Directory does not exist: ", datadir))
}
# Load mechanical shaker data for all five experiments
filename_five_experiments = paste0(datadir, "complete_data.RData") 
if (!file.exists(filename_five_experiments)) {
  stop(paste0("File does not exist: ", filename_five_experiments))
}
load(filename_five_experiments)
```

```{r Functions, include=FALSE}
#' plotSignals
#'
#' @description 'plotSignals' plots the signals in the data set
#'
#' @param outputdir A string that indicates the file path of folder (where the signal plot will be stored); not stored yet
#' @param data List of elements representing a data.frame for one signal, consisting of: 
#' \item{time}{Time stamps}
#' \item{x}{Acceleration signal of the x-axis}
#' \item{y}{Acceleration signal of the y-axis}
#' \item{z}{Acceleration signal of the z-axis}
#' \item{shaking_frequency}{Indicating the movement applied by the shaker machine in rpm}
#' \item{condition}{Indicating the attatchment of the accelerometers}
#' @param axes A vector of characters, one of c("x", "y", "z"), to indicate which axes are required for signal plotting
#' @return A list with the signal plots
#' 
#' @export
#' 
plotSignals <- function(outputdir, d, axes){
  plot_list <- list()
  for(signal in 1:length(d)) {
    file_name <- paste0("acceleration data: ", names(d)[signal])

    jpeg(paste0(outputdir, paste0(file_name, ".jpeg")), width=600, height=500, res=120) 
    colors  <- c( "x-axis" = "red", "y-axis" = "black", "z-axis" = "green")

    p <- ggplot2::ggplot()
    for(axis in 1:length(axes)){
      if(axes[axis] == "x"){
        p <- p + ggplot2::geom_point(data = d[[signal]], ggplot2::aes(x = time, y = x, color = "x-axis") , alpha = 0.2, size = 0.25)
      } else if(axes[axis] == "y"){
        p <- p + ggplot2::geom_point(data = d[[signal]], ggplot2::aes(x = time, y = y, color = "y-axis"), alpha = 0.2, size = 0.25) 
      } else if(axes[axis] == "z"){
        p <- p + ggplot2::geom_point(data = d[[signal]], ggplot2::aes(x = time, y = z, color = "z-axis"), alpha = 0.2, size = 0.25) }
    }
    
    
    p <- p + ggplot2::theme_bw() +
      ggplot2::labs(title = file_name, y = "acceleration (g)", color = "Axis") +
      ggplot2::scale_color_manual(values = colors)
    
    p
    dev.off()
    plot_list[[signal]] <- p
  }
  return(plot_list)
}
```

## Experiment 1 (E1): Low sampling rate and constant dynamic range (lrcr)
Visual inspection indicates no typical shaker movements for acceleration signals of aP_490 and aP_254.

```{r Signal plots ms_lrcr, include = FALSE}
index <- which(data$specifications$experiment == "ms_lrcr")
ms_lrcr <- list()
for (signal in 1:length(index)){
  ms_lrcr[[signal]] <- data$data[[index[signal]]]
}
names(ms_lrcr) <- data$specifications$label[index]
ms_lrcr <- Filter(Negate(is.null), ms_lrcr)

# Save a plot for each signal
plots.ms_lrcr <- plotSignals(outputdir = paste0(datadir, "plots"), d = ms_lrcr, axes = c("x", "y", "z"))
```

```{r Plot signals ms_lrcr, echo = FALSE}
# Save and print the plot for each signal
for(plot in 1:length(plots.ms_lrcr)){
  print(plots.ms_lrcr[[plot]])
}
rm(plot, plots.ms_lrcr)
```

## Experiment 2 (E2): High sampling rate and constant dynamic range (hrcr)
Note that we restarted this experiment, as we noticed that one GENEActiv device (GA_301) had come loose. MOX devices quit recording around 30 minutes before the end of the experiment.

```{r Signal plots ms_hrcr, include = FALSE}
index <- which(data$specifications$experiment == "ms_hrcr")
ms_hrcr <- list()
for (signal in 1:length(index)){
  ms_hrcr[[signal]] <- data$data[[index[signal]]]
}
names(ms_hrcr) <- data$specifications$label[index]
ms_hrcr <- Filter(Negate(is.null), ms_hrcr)

# Save a plot for each signal
plots.ms_hrcr <- plotSignals(outputdir = paste0(datadir, "plots"), d = ms_hrcr, axes = c("x", "y", "z"))
```

```{r Plot signals ms_hrcr, echo = FALSE}
# Save and print the plot for each signal
for(plot in 1:length(plots.ms_hrcr)){
  print(plots.ms_hrcr[[plot]])
}
rm(plot, plots.ms_hrcr)
```

## Experiment 3 (E3): Low sampling rate and mixed dynamic range (lrmr)
```{r Signal plots ms_lrmr, include = FALSE}
index <- which(data$specifications$experiment == "ms_lrmr")
ms_lrmr <- list()
for (signal in 1:length(index)){
  ms_lrmr[[signal]] <- data$data[[index[signal]]]
}
names(ms_lrmr) <- data$specifications$label[index]

# Save a plot for each signal
plots.ms_lrmr <- plotSignals(outputdir = paste0(datadir, "plots"), d = ms_lrmr, axes = c("x", "y", "z"))
```

```{r Plot signals ms_lmr, echo = FALSE}
# Save and print the plot for each signal
for(plot in 1:length(plots.ms_lrmr)){
  print(plots.ms_lrmr[[plot]])
}
rm(plot, plots.ms_lrmr)
```

## Experiment 4 (E4): High sampling rate and mixed dynamic range (hrmr)
```{r Signal plots ms_hrmr, include = FALSE}
index <- which(data$specifications$experiment == "ms_hrmr")
ms_hrmr <- list()
for (signal in 1:length(index)){
  ms_hrmr[[signal]] <- data$data[[index[signal]]]
}
names(ms_hrmr) <- data$specifications$label[index]

# Save a plot for each signal
plots.ms_hrmr <- plotSignals(outputdir = paste0(datadir, "plots"), d = ms_hrmr, axes = c("x", "y", "z"))
```

```{r Plot signals ms_hrmr, echo = FALSE}
# Save and print the plot for each signal
for(plot in 1:length(plots.ms_hrmr)){
  print(plots.ms_hrmr[[plot]])
}
rm(plot, plots.ms_hrmr)
```

## Experiment 5 (E5): Mixed sampling rate and constant dynamic range (mrcr)
```{r Signal plots ms_mrcr, include = FALSE}
index <- which(data$specifications$experiment == "ms_mrcr")
ms_mrcr <- list()
for (signal in 1:length(index)){
  ms_mrcr[[signal]] <- data$data[[index[signal]]]
}
names(ms_mrcr) <- data$specifications$label[index]
ms_mrcr <- Filter(Negate(is.null), ms_mrcr)

# Save a plot for each signal
plots.ms_mrcr <- plotSignals(outputdir = paste0(datadir, "plots"), d = ms_mrcr, axes = c("x", "y", "z"))
```

```{r Plot signals ms_mrcr, echo = FALSE}
# Save and print the plot for each signal
for(plot in 1:length(ms_mrcr)){
  print(plots.ms_mrcr[[plot]])
}
rm(plot, plots.ms_mrcr)
```

## Bag experiment 5 (ms_bag): High sampling rate and lack of standardized orientation (bag)

```{r Signal plots ms_bag, include = FALSE}
index <- which(data$specifications$experiment == "ms_bag")
ms_bag <- list()
for (signal in 1:length(index)){
  ms_bag[[signal]] <- data$data[[index[signal]]]
}
names(ms_bag) <- data$specifications$label[index]
ms_bag <- Filter(Negate(is.null), ms_bag)

# Save a plot for each signal
plots.ms_bag <- plotSignals(outputdir = paste0(datadir, "plots"), d = ms_bag, axes = c("x", "y", "z"))
```

```{r Plot signals ms_bag, echo = FALSE}
# Save and print the plot for each signal
for(plot in 1:length(plots.ms_bag)){
  print(plots.ms_bag[[plot]])
}
rm(plot, plots.ms_bag)
```
